---
title: Woodward (Kano) paradigm: Dwell time analysis
author: Christoph Voelter
date: 05/07/2021
output: html_document
---
notes: 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(summarytools)
library(exactRankTests)
library(ggpubr)
load("./workspaces/ET_dog_woodward_kano_paradigm.RData")

```

## Larger IAs
### End IP

```{r}
end.ip.orig.data <- read_delim("data/dog_woodward_paradigm_IA_report_largeAOI_video_IP_end.txt", na=".", delim="\t")

levels(as.factor(end.ip.orig.data$IA_LABEL))

demo.data <- read_csv("data/Woodward_Kano_ counterbalancing.csv")%>%
  mutate(EDF1=fct_recode(as.factor(EDF1), "Georgia_1"="Georgia1"))%>%
  separate(EDF1, c("subject", "num"), sep="_")



end.ip.data <- end.ip.orig.data %>%
  select(Session_Name_, IP_LABEL, IA_LABEL,IA_ID, Trial_Index_, condition,  target, target_location, session, trial_within_block, video_file,  video_latency, DATA_FILE,  IA_AREA, IA_AVERAGE_FIX_PUPIL_SIZE,  IA_DWELL_TIME, "IA_DWELL_TIME_%", IA_FIXATION_COUNT, IA_MAX_FIX_PUPIL_SIZE, INTEREST_AREA_FIXATION_SEQUENCE, TRIAL_DWELL_TIME, TRIAL_FIXATION_COUNT, TRIAL_IA_COUNT, TRIAL_TOTAL_VISITED_IA_COUNT, IA_FIRST_FIXATION_TIME )%>%
  mutate(Session_Name_=fct_recode(Session_Name_, Georgia_1="Georgia1",Georgia_2="Georgia2", Georgia_3="Georgia3",Georgia_4="Georgia4" ))%>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial=paste(condition, "_", trial_within_block))%>%
  full_join(demo.data)
  


end.ip.data.wider <- end.ip.data %>%
  select(
    subject,
    session.num,
    condition_trial,
    IP_LABEL,
    IA_LABEL,
    condition,
    condition_first,
    target,
    target_location,
    session,
    trial_within_block,
    video_file,
    video_latency,
    DATA_FILE,
    IA_DWELL_TIME
  ) %>%
  pivot_wider(names_from = IA_LABEL, values_from = IA_DWELL_TIME) %>%
  mutate(
    object_congruent_dwell_time = as.numeric(ifelse(
      target_location == "up",
      lower_IA_large,
      ifelse(target_location == "down", upper_IA_large, "")
    )),
    object_IA_dwell_time = lower_IA_large + upper_IA_large,
    object_incongruent_dwell_time = as.numeric(ifelse(
      target_location == "up",
      upper_IA_large,
      ifelse(target_location == "down", lower_IA_large, "")
    ))
  ) %>%
  select(-lower_IA_large,-upper_IA_large) %>%
  pivot_longer(
    cols = c(object_congruent_dwell_time, object_incongruent_dwell_time),
    names_to = "object_congruent",
    values_to = "dwell_time"
  ) %>%
  mutate(
    proportion.object.congruent.dwell.time = dwell_time / object_IA_dwell_time,
    proportion.object.congruent.dwell.time_noNA = as.numeric(ifelse(
      is.na(proportion.object.congruent.dwell.time),
      0,
      proportion.object.congruent.dwell.time
    )
    ))


end.ip.first.look <- end.ip.data %>%
  select(
    subject,
    session.num,
    condition_trial,
    IP_LABEL,
    IA_LABEL,
    condition,
    condition_first,
    target,
    target_location,
    session,
    trial_within_block,
    video_file,
    video_latency,
    DATA_FILE,
    IA_FIRST_FIXATION_TIME
  ) %>%
  pivot_wider(names_from = IA_LABEL, values_from = IA_FIRST_FIXATION_TIME) %>%
  mutate(
    upper_IA_large = as.numeric(upper_IA_large),
    lower_IA_large = as.numeric(lower_IA_large)
  ) %>%
  mutate(first_look = ifelse((
    upper_IA_large < lower_IA_large |
      (!is.na(upper_IA_large) & is.na(lower_IA_large))
  ), "upper", "")) %>%
  mutate(first_look2 = ifelse((
    lower_IA_large < upper_IA_large |
      (!is.na(lower_IA_large) & is.na(upper_IA_large))
  ), "lower", "")) %>%
  mutate(first_look3 = ifelse(
    !is.na(first_look),
    first_look,
    ifelse(!is.na(first_look2), first_look2, NA)
  )) %>%
  mutate(object_congruent_first_look = as.numeric(ifelse((target_location ==
                                                            "up" &
                                                            first_look3 == "lower"),
                                                         1,
                                                         ifelse((target_location == "down" &
                                                                   first_look3 == "upper"),
                                                                1,
                                                                ifelse((target_location == "up" &
                                                                          first_look3 == "upper"),
                                                                       0,
                                                                       ifelse((
                                                                         target_location == "down" & first_look3 == "lower"
                                                                       ), 0, NA)
                                                                )
                                                         )
  )),
  object_congruent_first_look_noNA = as.numeric(ifelse(is.na(first_look3),0, object_congruent_first_look )))

#checks:
view(dfSummary(end.ip.data))

table(end.ip.data$trial_within_block, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$subject)
table(end.ip.data$session.num, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$TRIAL_TOTAL_VISITED_IA_COUNT)

```


* Absolute dwell time
```{r}

test.IP.end.data <- end.ip.data %>%
  group_by(subject, condition, target_location, IA_LABEL)%>%
  summarise(mean_dwell=mean(IA_DWELL_TIME))



ggplot(data = test.IP.end.data%>%filter(condition=="claw"), aes(x=IA_LABEL, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  #geom_point(aes(x=first_target_loc, y=IA_DWELL_TIME, group=IA_LABEL),position_dodge(), alpha=0.5)+
  facet_grid(~target_location)+
  ylab("Dwell time")+xlab("Condition")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )


ggplot(data = test.IP.end.data%>%filter(condition=="hand"), aes(x=IA_LABEL, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  #geom_point(aes(x=first_target_loc, y=IA_DWELL_TIME, group=IA_LABEL),position_dodge(), alpha=0.5)+
  facet_grid(~target_location)+
  ylab("Dwell time")+xlab("Condition")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```
* Proportion dwell time
```{r}

test.IP.end.data.object.congruent <- end.ip.data.wider %>%
  group_by(subject, condition, object_congruent)%>%
  summarise(mean_dwell=mean(dwell_time), mean_proportion_dwell=mean(proportion.object.congruent.dwell.time, na.rm = TRUE))



ggplot(data = test.IP.end.data.object.congruent, aes(x=object_congruent, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  geom_point(alpha=0.3)+
  geom_line(aes(group=subject), alpha=0.3)+
  facet_wrap(~condition)+
  ylab("Dwell time")+xlab("Object congruency")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )


ggplot(data = test.IP.end.data.object.congruent, aes(x=object_congruent, y=mean_proportion_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  geom_point(alpha=0.3)+
  geom_line(aes(group=subject), alpha=0.3)+
  facet_wrap(~condition)+
  ylab("Proportion dwell time")+xlab("Object congruency")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 0.8 )
```

* T-test comparing dwell times between object congruent and incongruent IAs
```{r}
test.IP.end.data.object.congruent.hand<-test.IP.end.data.object.congruent%>%
  filter(condition=="hand")

t.test(test.IP.end.data.object.congruent.hand$mean_dwell[test.IP.end.data.object.congruent.hand$object_congruent=="object_congruent_dwell_time"], test.IP.end.data.object.congruent.hand$mean_dwell[test.IP.end.data.object.congruent.hand$object_congruent=="object_incongruent_dwell_time"], paired=TRUE)
```


#### Beta GLMM

```{r}

model.data.end.IP.largerIA<-end.ip.data.wider%>%
  filter(object_congruent=="object_congruent_dwell_time")%>%
  mutate(z.trial=scale(trial_within_block, center = TRUE, scale=TRUE), # trial within condition
         condition.c=scale(as.numeric(as.factor(condition)), center = TRUE, scale=FALSE))

summary(model.data.end.IP.largerIA)
```

```{r}
model.data.end.IP.largerIA$prop.dwell.scaled <-
  (
    model.data.end.IP.largerIA$proportion.object.congruent.dwell.time_noNA *
      (
        length(
          model.data.end.IP.largerIA$proportion.object.congruent.dwell.time_noNA
        ) - 1
      ) + 0.5
  ) / length(model.data.end.IP.largerIA$proportion.object.congruent.dwell.time_noNA)#transform DV to exclude 0 and 1s

library(glmmTMB)

m1.endIP.largeIA<-glmmTMB(prop.dwell.scaled ~ condition + z.trial + condition_first +  target_location+(1+condition.c+z.trial||subject), family=beta_family, data=model.data.end.IP.largerIA,  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))

summary(m1.endIP.largeIA)
```


```{r}
drop1_m1<-drop1(m1.endIP.largeIA, test="Chisq")
drop1_m1
```

```{r}


m2.endIP.largeIA<-glmmTMB(prop.dwell.scaled ~ condition*target_location+z.trial +(1+condition.c+z.trial||subject), family=beta_family, data=model.data.end.IP.largerIA,  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))

```

```{r}
drop1_m2<-drop1(m2.endIP.largeIA, test="Chisq")
drop1_m2
```

The results of the dwell time analysis of the end IP show no difference between the hand or claw condition with respect to the looking times to the grasped object. Dogs showed a preference to look at the upper location (evidenced by a significant target location effect).

### Intercept only model for proportion congruent object dwell time
```{r}
model.data.end.IP.largerIA$prop.dwell.scaled2 <-
  (
    model.data.end.IP.largerIA$proportion.object.congruent.dwell.time *
      (
        length(
          model.data.end.IP.largerIA$proportion.object.congruent.dwell.time
        ) - 1
      ) + 0.5
  ) / length(model.data.end.IP.largerIA$proportion.object.congruent.dwell.time)#transform DV to exclude 0 and 1s

m1.endIP.largeIA_interceptonly<-glmmTMB(prop.dwell.scaled2 ~  z.trial+(1+z.trial||subject), family=beta_family, data=model.data.end.IP.largerIA,  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))
summary(m1.endIP.largeIA_interceptonly)
```

### first look analysis (binary)
```{r}
model.data.end.IP.first.looks<-end.ip.first.look%>%
  mutate(z.trial=scale(trial_within_block, center = TRUE, scale=TRUE),
         condition.c=scale(as.numeric(as.factor(condition)), center = TRUE, scale=FALSE))

summary(model.data.end.IP.first.looks)


library(lme4)

m1.endIP.first_look<-glmer(object_congruent_first_look_noNA ~ condition + z.trial + condition_first+ target_location+(1+condition.c+z.trial|subject), family=binomial, data=model.data.end.IP.first.looks, control=glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000000)))

summary(m1.endIP.first_look)
```


```{r}
drop1_m1.first.look<-drop1(m1.endIP.first_look, test="Chisq")
drop1_m1.first.look
```

### Intercept only model for first look
```{r}
m1.endIP.first_look_interceptonly<-glmer(object_congruent_first_look ~ 1+z.trial+(1+z.trial|subject), family=binomial, data=model.data.end.IP.first.looks, control=glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000000)))

summary(m1.endIP.first_look_interceptonly)
```





### End IP (longer)

```{r}
end.ip.long.orig.data <- read_delim("data/dog_woodward_paradigm_IA_report_largeAOI_video_IP_end_long.txt", na=".", delim="\t")

demo.data <- read_csv("data/Woodward_Kano_ counterbalancing.csv")%>%
  mutate(EDF1=fct_recode(as.factor(EDF1), "Georgia_1"="Georgia1"))%>%
  separate(EDF1, c("subject", "num"), sep="_")


end.ip.long.data <- end.ip.long.orig.data %>%
  select(Session_Name_, IP_LABEL, IA_LABEL,IA_ID, Trial_Index_, condition, target, target_location, session, trial_within_block, video_file,  video_latency, DATA_FILE,  IA_AREA, IA_AVERAGE_FIX_PUPIL_SIZE,  IA_DWELL_TIME, "IA_DWELL_TIME_%", IA_FIRST_FIXATION_TIME, IA_FIXATION_COUNT, IA_MAX_FIX_PUPIL_SIZE, INTEREST_AREA_FIXATION_SEQUENCE, TRIAL_DWELL_TIME, TRIAL_FIXATION_COUNT, TRIAL_IA_COUNT, TRIAL_TOTAL_VISITED_IA_COUNT )%>%
 #  filter(!(DATA_FILE=="Ace_2.edf" & Trial_Index_!=4))%>%
#  mutate(trial=ifelse(DATA_FILE=="Ace_2.edf" & Trial_Index_==4, 5, Trial_Index_))%>% #recoding of trial number to account for a dog that did not complete his first session, session was repeated with this dog
  mutate(Session_Name_=fct_recode(Session_Name_, Georgia_1="Georgia1",Georgia_2="Georgia2", Georgia_3="Georgia3",Georgia_4="Georgia4" ))%>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial=paste(condition, "_", trial_within_block))%>%
  full_join(demo.data)
  


end.ip.long.data.wider<-end.ip.long.data%>%
  select(subject, session.num,condition_trial, IP_LABEL, IA_LABEL, condition, target, target_location, session, trial_within_block, video_file,  video_latency, DATA_FILE,  IA_DWELL_TIME)%>%
  pivot_wider(names_from = IA_LABEL, values_from=IA_DWELL_TIME)%>%
  mutate(object_congruent_dwell_time=as.numeric(ifelse(target_location=="up",lower_IA_large, ifelse(target_location=="down",upper_IA_large, "") )), object_incongruent_dwell_time=as.numeric(ifelse(target_location=="up",upper_IA_large, ifelse(target_location=="down",lower_IA_large, "") )))%>%
  select(-lower_IA_large, -upper_IA_large)%>%
  pivot_longer(cols=c(object_congruent_dwell_time, object_incongruent_dwell_time), names_to = "object_congruent", values_to = "dwell_time")


#checks:
view(dfSummary(end.ip.data))

table(end.ip.data$trial_within_block, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$subject)
table(end.ip.data$session.num, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$TRIAL_TOTAL_VISITED_IA_COUNT)

```


```{r}

test.IP.end.long.data <- end.ip.long.data %>%
  group_by(subject, condition, target_location, IA_LABEL)%>%
  summarise(mean_dwell=mean(IA_DWELL_TIME))



ggplot(data = test.IP.end.long.data, aes(x=IA_LABEL, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  #geom_point(aes(x=first_target_loc, y=IA_DWELL_TIME, group=IA_LABEL),position_dodge(), alpha=0.5)+
  facet_grid(condition~target_location)+
  ylab("Dwell time")+xlab("Condition")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```
Object congruent looks
```{r}

test.IP.end.long.data.object.congruent <- end.ip.long.data.wider %>%
  group_by(subject, condition, object_congruent)%>%
  summarise(mean_dwell=mean(dwell_time))



ggplot(data = test.IP.end.long.data.object.congruent, aes(x=object_congruent, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  geom_point(alpha=0.3)+
  geom_line(aes(group=subject), alpha=0.3)+
  facet_wrap(~condition)+
  ylab("Dwell time")+xlab("Object congruency")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```






### Trial report

```{r}

trialreport.end.ip.orig.data <- read_delim("data/dog_woodward_paradigm_trial_report_video_IP_end.txt", na=".", delim="\t")



```







-repear IA analysis with larger areas of interest
- setup a model for the dwell time (proportion of correct AOI)
- first look
- pupil size







############################################################

```{r}
#first test trial
test.IP.end.IA.endpos.trial1 <- end.ip.data %>%
  filter(condition!="hab", IA_LABEL=="IA_end_position", trial_within_block==4)


t.test(test.IP.end.IA.endpos.trial1$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="exp"], test.IP.end.IA.endpos.trial1$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="con"], paired=TRUE)

#second test trial
test.IP.end.IA.endpos.trial2 <- end.ip.data %>%
  filter(condition!="hab", IA_LABEL=="IA_end_position", trial==5)


t.test(test.IP.end.IA.endpos.trial2$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="exp"], test.IP.end.IA.endpos.trial2$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="con"], paired=TRUE)


```


### Onscreen Dwell time (while videos is playing)

```{r}
video_playing_data<-read.delim("data/ET_VoE_occlusion2screens_IA_report_whole_screen_video_playing_IP.txt",na=".", sep = "\t")%>%
  select(Session_Name_, IP_LABEL, IA_LABEL,IA_ID, Trial_Index_, condition, phase, session,  video_file, video_id, video_latency, DATA_FILE,  IA_AREA, IA_AVERAGE_FIX_PUPIL_SIZE,  IA_DWELL_TIME, "IA_DWELL_TIME", IA_FIXATION_COUNT, IA_MAX_FIX_PUPIL_SIZE, INTEREST_AREA_FIXATION_SEQUENCE, TRIAL_DWELL_TIME, TRIAL_FIXATION_COUNT, TRIAL_IA_COUNT, TRIAL_TOTAL_VISITED_IA_COUNT )%>%
   filter(!(DATA_FILE=="Ace_2.edf" & Trial_Index_!=4))%>%
  mutate(trial=ifelse(DATA_FILE=="Ace_2.edf" & Trial_Index_==4, 5, Trial_Index_))%>% #recoding of trial number to account for a dog that did not complete his first session, session was repeated with this dog
  mutate(Session_Name_=fct_recode(Session_Name_, Georgia_1="Georgia1",Georgia_2="Georgia2", Cheynna_1="Cheynna1", Cheynna_2="Cheynna2" ))%>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial=paste(condition, "_", trial))%>%
  full_join(demo.data)
 # rename(trial="Trial_Index_")
  

#checks:
view(dfSummary(video_playing_data))


on_screen_data<-video_playing_data%>%
  filter(IA_LABEL=="screen_aoi")%>%
  mutate(prop_looking=IA_DWELL_TIME/5000)


on_screen_data%>%group_by(condition)%>%
  summarise(mean=mean(prop_looking, na.rm = TRUE), median=median(prop_looking, na.rm = TRUE), se=sd(prop_looking, na.rm = TRUE)/sqrt(length(prop_looking)), min=min(prop_looking, na.rm = TRUE), max=max(prop_looking, na.rm = TRUE))

```

onscreen dwell-time: comparison between first and last familiarisation trial.
```{r}
library(broom)
on_screen_data%>%filter(phase=="hab1" | phase=="hab3")%>%
  group_by(subject, phase)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ phase, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```


onscreen dwell-time: comparison between test trials
* first test trial
```{r}
on_screen_data%>%filter(condition!="hab", trial==4)%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ condition, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```

* second test trial
```{r}
on_screen_data%>%filter(condition!="hab", trial==5)%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ condition, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```

Plot onscreen dwell time in test trials
```{r}
xx<-on_screen_data%>%filter(condition!="hab")%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking))%>%
  ungroup()

ggplot(data=xx, aes(x=condition, y=mean))+
  geom_boxplot()
```