---
title: Woodward 2022: Dwell time analysis
author: Christoph Voelter
date: 
output: html_document
---
notes: 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#rm(list=ls())
#remove(list=c("first4s.data_short"))
library(tidyverse)
library(summarytools)
library(lme4)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
source("./functions/glmmTMB_stability.r")
source("./functions/boot_glmmTMB.r")
source("./functions/extract_ranef_gmmTMB.r")
load("workspaces/AL_test_first4s.RData")
```

```{r}
save.image("workspaces/AL_test_first4s.RData")
```


```{r}
demo.data <- read_csv("data/Woodward_2022_counterbalancing.csv") %>%
  mutate(
    EDF1 = fct_recode(
      as.factor(EDF1),
      "Georgia_1" = "Georgia1",
    )
  ) %>%
  separate(EDF1, c("subject", "num"), sep = "_")
```

```{r}
first4s.data <-
  read_delim(
    "data/Woodward2022_IA_Report_IP_test_first_4_sec.txt",
    na = ".",
    delim = "\t"
  ) %>%
  dplyr::select(
    Session_Name_,
    IP_LABEL,
    IA_LABEL,
    IA_ID,
    Trial_Index_,
    condition,
    dog_session,
    sub_id,
    trial_number,
    trial_type,
    session_number,
    video_file,
    DATA_FILE,
    IA_AREA,
    IA_AVERAGE_FIX_PUPIL_SIZE,
    IA_DWELL_TIME,
    "IA_DWELL_TIME_%",
    IA_FIXATION_COUNT,
    "IA_FIXATION_%",
    IA_FIRST_FIXATION_INDEX, #Ordinal sequence of the first fixation that       #was within the current interest area.
    IA_FIRST_FIXATION_VISITED_IA_COUNT, #number of different interest areas
    #visited so far before the first fixation is made to the
    #current interest area.
    IA_MAX_FIX_PUPIL_SIZE,
    IA_FIRST_FIXATION_TIME,
    INTEREST_AREA_FIXATION_SEQUENCE,
    TRIAL_DWELL_TIME,
    TRIAL_FIXATION_COUNT,
    TRIAL_IA_COUNT,
    TRIAL_TOTAL_VISITED_IA_COUNT,
    IP_START_TIME,
    TRIAL_START_TIME
  ) %>%
  filter(DATA_FILE != "Melody_1.edf", DATA_FILE != "Mia_5.edf") %>%
  filter(DATA_FILE != "Timo_5.edf", DATA_FILE != "Joker_5.edf") %>% # backup recordings in case the first sessions are below 70% onscreen looking
  #filter(!(DATA_FILE == "Amy_4.edf" & Trial_Index_ != 1)) %>%
  #mutate(session.num = as.numeric(session.num)) %>%
  #mutate(session.num.corrected = ifelse(DATA_FILE == "Amy_4.edf" | DATA_FILE == "Amy_5.edf" | DATA_FILE == "George_5.edf" | DATA_FILE == "Melody_2.edf" | DATA_FILE == #"Melody_3.edf" | DATA_FILE == "Melody_4.edf", session.num-1,
#                                        ifelse(DATA_FILE == "Melody_7.edf" | DATA_FILE == "Mia_6.edf", session.num-3, session.num))) %>% #recoding of session number to account for dogs with aborted sessions
  mutate(
    Session_Name_ = fct_recode(
      Session_Name_,
      Georgia_1 = "Georgia1",
      Georgia_2 = "Georgia2",
      Georgia_3 = "Georgia3",
      Georgia_4 = "Georgia4"
    )
  ) %>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial = paste(condition, "_", Trial_Index_)) %>%
  full_join(demo.data) %>%
  mutate(IA_LABEL2 = IA_LABEL, trial_type2 = trial_type) %>%
  separate(
    IA_LABEL2,
    into = c("ia_object", "ia_side", "ia_label_del", "ia_label_del2"),
    sep = "_"
  ) %>%
  separate(
    trial_type2,
    into = c("trialtype_agent", "trialtype_object", "trialtype_side"),
    sep = "_"
  ) %>%
  filter(ia_object == "elephant" | ia_object == "ball") %>%
  mutate(
    IA_fam_object = ifelse(
      condition == "new_goal" &
        ia_object == trialtype_object,
      "new_goal_old_side",
      ifelse(
        condition == "new_goal" &
          ia_object != trialtype_object,
        "old_goal_new_side",
        ifelse(
          condition == "new_side" &
            ia_object != trialtype_object,
          "new_goal_old_side",
          ifelse(
            condition == "new_side" &
              ia_object == trialtype_object,
            "old_goal_new_side",
            ""
          )
        )
      )
    )
  ) %>%
  mutate(attention_getter_duration = IP_START_TIME - TRIAL_START_TIME)

first4s.data %>% select(subject, condition_trial, trial_type, IA_LABEL, IA_fam_object) 
#IA_fam_object=what the AOI looked at contains
first4s.data %>% select(subject, condition, trial_type, IA_LABEL, IA_fam_object,IA_FIRST_FIXATION_INDEX,IA_FIRST_FIXATION_VISITED_IA_COUNT) 


levels(as.factor(first4s.data$DATA_FILE))
levels(as.factor(first4s.data$condition))
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$Trial_Index_)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$Trial_Index_)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$session.num)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$session_number)
```

```{r}

first4s.data.first.fix <- first4s.data %>%
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  select(subject, session_number, IA_fam_object , condition, ,trial_type, trialtype_agent, agent_target_side_first, IA_FIRST_FIXATION_TIME)%>%
  pivot_wider(names_from = IA_fam_object, values_from=IA_FIRST_FIXATION_TIME)%>%
  mutate(old_goal_new_side = replace_na(old_goal_new_side, 99999999), new_goal_old_side = replace_na(new_goal_old_side, 99999999))%>%
  mutate(first_fix_old_goal=as.numeric(ifelse(old_goal_new_side<new_goal_old_side, 1, 0)),
         first_fix_old_goal_withoutNA = as.numeric(ifelse(old_goal_new_side==new_goal_old_side, NA,
                                                          ifelse(old_goal_new_side<new_goal_old_side, 1, 0))),
         first_fix_old_side=as.numeric(ifelse(old_goal_new_side>new_goal_old_side, 1, 0)),
         first_fix_old_side_withoutNA = as.numeric(ifelse(old_goal_new_side==new_goal_old_side, NA,
                                                          ifelse(old_goal_new_side>new_goal_old_side, 1, 0)))) %>%
    separate(
    agent_target_side_first,
    into = c("agent_first", "object_first", "side_first"), #shown to the dog in the first session
    sep = "_"
  ) %>%
  mutate(session_number = as.numeric(session_number), session_withinagent = as.numeric(ifelse(session_number >=3, session_number-2, session_number)))



# find the first looks to the old goal or old side. If there were first looks to one of the two IAa in both sessions we retained the first session value
first4s.data.first.fix.long <- first4s.data.first.fix %>%
  filter(first_fix_old_goal == 1 | first_fix_old_side == 1) %>%
  select(
    subject,
    trialtype_agent,
    session_withinagent,
    first_fix_old_goal,
    first_fix_old_side
  ) %>%
  pivot_longer(
    cols = c(first_fix_old_goal, first_fix_old_side),
    names_to = "IA",
    values_to = "looked"
  ) %>%
  pivot_wider(names_from = session_withinagent, values_from = looked) %>%
  rename("first.s" = "1", "second.s" = "2") %>%
  mutate(first.s = replace_na(first.s, 0), second.s = replace_na(second.s, 0))%>%
  mutate(look_overall = as.numeric(ifelse(first.s == 1, 1, ifelse(second.s == 1, 1, 0))))%>%
  filter(look_overall ==1)


table(first4s.data.first.fix.long$trialtype_agent, first4s.data.first.fix.long$IA)
mcnemar.test(table(first4s.data.first.fix.long$trialtype_agent, first4s.data.first.fix.long$IA))


```


## first fixation (binary model and no fixations are counted as 0)
```{r}
table(first4s.data.first.fix$subject, first4s.data.first.fix$session_number)
table(first4s.data.first.fix$subject, first4s.data.first.fix$session_withinagent)

first4s.data.first.fix$z.session <-
  as.vector(scale(
    as.numeric(first4s.data.first.fix$session_number),
    center = TRUE,
    scale = TRUE
  ))

first4s.data.first.fix$z.session_within <-
  as.vector(scale(
    as.numeric(first4s.data.first.fix$session_withinagent),
    center = TRUE,
    scale = TRUE
  ))

first4s.data.first.fix$trialtype_agent <-
  as.factor(first4s.data.first.fix$trialtype_agent)
levels(first4s.data.first.fix$trialtype_agent)

first4s.data.first.fix$trialtype_agent.c = as.vector(scale(
  as.numeric(
    first4s.data.first.fix$trialtype_agent == levels(as.factor(first4s.data.first.fix$trialtype_agent))[2]
  ),
  center = TRUE,
  scale = FALSE
))

first4s.data.first.fix$agent_first.c = as.vector(scale(
  as.numeric(first4s.data.first.fix$agent_first == levels(
    as.factor(first4s.data.first.fix$agent_first)
  )[2]),
  center = TRUE,
  scale = FALSE
))
levels(as.factor(first4s.data.first.fix$z.session))
hist(first4s.data.first.fix$first_fix_old_goal)
```

fit model
```{r}
mm1_first_fix=glmer(first_fix_old_goal ~ trialtype_agent+z.session_within+agent_first +
            (1+trialtype_agent.c + z.session_within|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#prune random slope structure due to convergence warning (i.e, exclude correlation between random icpt and slopes, remove random slope of session number)
mm1_first_fix2=glmer(first_fix_old_goal ~ trialtype_agent+z.session_within+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(mm1_first_fix2)
```


```{r}

mm1_first_fix2_drop1 <- drop1(mm1_first_fix2, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1)
```


Check for collinearity
```{r}
library(car)
xx=lm(first_fix_old_goal ~ trialtype_agent+z.session_within+agent_first, data=first4s.data.first.fix)
vif(xx)
```
#### Confidence intervals of the estimates
```{r}
mm1_first_fix2.ci=boot.glmm.pred(model.res=mm1_first_fix2, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)
mm1_first_fix2.ci$ci.estimates

```

#### model stability
```{r}
m.stab.mm1_first_fix2 <- glmm.model.stab(model.res = mm1_first_fix2)
m.stab.mm1_first_fix2$detailed$warnings
m.stab.mm1_first_fix2_output <- as.data.frame(round(m.stab.mm1_first_fix2$summary[, -1], 3))
m.stab.mm1_first_fix2_output 
png("graphs/mm1_first_fix2_stability_plot.png")
m.stab.plot(round(m.stab.mm1_first_fix2$summary[, -1], 3))
dev.off()
m.stab.plot(round(m.stab.mm1_first_fix2$summary[, -1], 3))
# The model appeared to be stable with respect to the fixed effects (see full_model_stability_plot).

```


#### output table


```{r}
mm1_first_fix2_output_table <-
  bind_cols(as.data.frame(summary(mm1_first_fix2)$coeff),
            mm1_first_fix2_drop1,
            mm1_first_fix2.ci$ci.estimates,
           m.stab.mm1_first_fix2_output[1:4,]) %>%
  select(
    Estimate,
    SE = `Std. Error`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`,
    min_stab = min,
    max_stab = max
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  mutate(across(.cols = c(z_wald:max_stab), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_first_fix2_output_table, file = "saves/mm1_first_fix2_output_table.csv")
```

## Plot showing the object looks in the first 4 s of the test trials (including dogs that did not look at any of the objects in a given trial)
```{r}
first4s.data.first.fix.nolooked.included.agg<-first4s.data.first.fix %>%
  group_by(trialtype_agent, session_withinagent) %>%
  summarise(first_fix_old_goal_mean = mean(first_fix_old_goal, na.rm=TRUE), 
            first_fix_old_goal_se = sd(first_fix_old_goal, na.rm=TRUE) / sqrt(length(first_fix_old_goal)),
            #number of dogs that anticipated the old goal object:
            first_fix_old_goal_sum = sum(first_fix_old_goal==1, na.rm=TRUE), 
            #number of dogs that looked at at least one of the objects:
            first_fix_old_goal_count = sum(!is.na(first_fix_old_goal)),
            first_fix_old_side_mean = mean(first_fix_old_side, na.rm=TRUE),
            first_fix_old_side_se = sd(first_fix_old_side, na.rm=TRUE) / sqrt(length(first_fix_old_side)),
            #number of dogs that anticipated the old side object:
            first_fix_old_side_sum = sum(first_fix_old_side==1, na.rm=TRUE), 
            #number of dogs that looked at at least one of the objects:
            first_fix_old_side_count = sum(!is.na(first_fix_old_side)))



al_plot_NAs_included<-ggplot(first4s.data.first.fix.nolooked.included.agg, aes(x = trialtype_agent, y = first_fix_old_goal_mean)) +
  geom_bar(stat="identity") + 
  facet_wrap(~session_withinagent)+
  ylim(0,1)+
  ylab("Prop. dogs fixating the old goal first")+
  xlab("Agent")+
  theme(axis.text = element_text(size = 14))+  #axis text size 
  theme(axis.title = element_text(size = 16))
```



```{r}
ggsave(filename="graphs/AL_NAs_included.png",
  plot = al_plot_NAs_included,
  width=8.5,
  height=5,
  #scale=0.55,
  dpi=1200)
```


```{r}
first4s.data.first.fix.nolooked.included.se.long <- first4s.data.first.fix.nolooked.included.agg %>%
  select(trialtype_agent, session_withinagent, first_fix_old_goal = first_fix_old_goal_se, first_fix_old_side = first_fix_old_side_se) %>%
  pivot_longer(cols = c(first_fix_old_goal, first_fix_old_side), names_to = "AOI", values_to = "se.dogs")

first4s.data.first.fix.nolooked.included.agg.long <- first4s.data.first.fix.nolooked.included.agg %>%
  select(trialtype_agent, session_withinagent, first_fix_old_goal = first_fix_old_goal_mean, first_fix_old_side = first_fix_old_side_mean) %>%
  pivot_longer(cols = c(first_fix_old_goal, first_fix_old_side), names_to = "AOI", values_to = "prop.dogs") %>%
  full_join(first4s.data.first.fix.nolooked.included.se.long)

al_plot_NAs_included2<-ggplot(first4s.data.first.fix.nolooked.included.agg.long, aes(x = trialtype_agent, y = prop.dogs, fill = AOI)) +
  geom_bar(stat="identity", position = position_dodge(0.9)) + 
  geom_errorbar(aes(ymin = prop.dogs - se.dogs, ymax = prop.dogs + se.dogs), width = 0.2, position = position_dodge(0.9))+
  facet_wrap(~session_withinagent)+
  ylim(0,1)+
  ylab("Prop. dogs")+
  xlab("Agent")+
  theme(axis.text = element_text(size = 14))+  #axis text size 
  theme(axis.title = element_text(size = 16)) +
  theme_bw() +
  scale_fill_manual(values = c("darkorange", "dodgerblue"), labels = c(first_fix_old_goal = "old goal", first_fix_old_side = "old side"))

al_plot_NAs_included2
```
```{r}
ggsave(filename="graphs/AL_NAs_excluded_AOIs.png",
  plot = al_plot_NAs_included2,
  width=8.5,
  height=5,
  scale=0.55,
  dpi=1200)
```

## first fixation (binary model and no fixations are treated as NA)
```{r}
hist(first4s.data.first.fix$first_fix_old_goal_withoutNA)
```

fit model
```{r}
mm1_first_fix_withoutNA=glmer(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session_within+agent_first +
            (1+trialtype_agent.c + z.session_within|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
#prune random slope structure due to convergence warning
mm1_first_fix_withoutNA2=glmer(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session_within+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),#removed the correlation between random intercept and random slope
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
#model has unusually large predictor estimates and variance of the random slopes
summary(mm1_first_fix_withoutNA2)
mm1_first_fix_withoutNA2_drop1<-drop1(mm1_first_fix_withoutNA2, test="Chisq")
```

pruned model (random slope of z.session removed)
```{r}
mm1_first_fix_withoutNA2_pruned=glmer(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),#removed the correlation between random intercept and random slope and the random slope of session
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
# estimates and random slope variance look better (values are smaller)
#null model because drop 1 does not converge
mm1_first_fix_withoutNA2_pruned_null=glmer(first_fix_old_goal_withoutNA ~ z.session+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),#removed the correlation between random intercept and random slope
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
anova(mm1_first_fix_withoutNA2_pruned, mm1_first_fix_withoutNA2_pruned_null, test = "Chisq")
#--> agent not significant 
summary(mm1_first_fix_withoutNA2_pruned)
drop1(mm1_first_fix_withoutNA2_pruned, test = "Chisq") #does not converge
```



Check for collinearity - model without NAs
```{r}
library(car)
xx=lm(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session+agent_first, data=first4s.data.first.fix)
vif(xx)
```
#### Confidence intervals of the estimates
```{r}
mm1_first_fix2_withoutNAs.ci=boot.glmm.pred(model.res=mm1_first_fix_withoutNA2, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)
mm1_first_fix2_withoutNAs.ci$ci.estimates

```

#### model stability
```{r}
m.stab.mm1_first_fix2_withoutNAs <- glmm.model.stab(model.res = mm1_first_fix_withoutNA2)
m.stab.mm1_first_fix2_withoutNAs$detailed$warnings #models failed to converge
m.stab.mm1_first_fix2_withoutNAs_output <- as.data.frame(round(m.stab.mm1_first_fix2_withoutNAs$summary[, -1], 3))
m.stab.mm1_first_fix2_withoutNAs_output 
png("graphs/mm1_first_fix2_withoutNAs_stability_plot.png")
m.stab.plot(round(m.stab.mm1_first_fix2_withoutNAs$summary[, -1], 3))
#dev.off()
# The estimates do not look reasonable
```


#### output table


```{r}
mm1_first_fix2_withoutNAs_output_table <-
  bind_cols(as.data.frame(summary(mm1_first_fix_withoutNA2)$coeff),
            mm1_first_fix_withoutNA2_drop1,
            mm1_first_fix2_withoutNAs.ci$ci.estimates,
            m.stab.mm1_first_fix2_withoutNAs_output[1:4,]) %>%
  select(
    Estimate,
    SE = `Std. Error`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`,
    min_stab = min,
    max_stab = max
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  mutate(across(.cols = c(z_wald:max_stab), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_first_fix2_withoutNAs_output_table, file = "saves/mm1_first_fix2_withoutNAs_output_table.csv")
```

## Plot showing the object looks in the first 4 s of the test trials (without dogs that did not look not at any of the objects in a given trial)
```{r}
first4s.data.first.fix.agg<-first4s.data.first.fix %>%
  group_by(trialtype_agent, agent_first, session_withinagent) %>%
  summarise(first_fix_old_goal_mean = mean(first_fix_old_goal_withoutNA, na.rm=TRUE), 
            #number of dogs that anticipated the old goal object:
            first_fix_old_goal_sum = sum(first_fix_old_goal_withoutNA==1, na.rm=TRUE),   
            # number of dogs that looked at at least one of the objects:
            first_fix_old_goal_count = sum(!is.na(first_fix_old_goal_withoutNA)))
```


```{r}
first4s.data.first.fix.agg2<-first4s.data.first.fix %>%
  group_by(trialtype_agent) %>%
  summarise(first_fix_old_goal_mean = mean(first_fix_old_goal_withoutNA, na.rm=TRUE), 
            first_fix_old_goal_sum = sum(first_fix_old_goal_withoutNA==1, na.rm=TRUE), #number of dogs that anticipated the old goal object
            first_fix_old_side_sum = sum(first_fix_old_goal_withoutNA==0, na.rm=TRUE), #number of dogs that anticipated the old side object
            first_fix_old_goal_count = sum(!is.na(first_fix_old_goal_withoutNA)))#number of dogs that looked at at least one of the objects

#binomial test for
#human agent
binom.test(x = first4s.data.first.fix.agg2$first_fix_old_goal_sum[1] , first4s.data.first.fix.agg2$first_fix_old_goal_count[1] )
#and for inanimate agent
binom.test(x = first4s.data.first.fix.agg2$first_fix_old_goal_sum[2] , first4s.data.first.fix.agg2$first_fix_old_goal_count[2] )
```

```{r}

AL_plot_NAs_excluded<-ggplot(first4s.data.first.fix.agg, aes(x = trialtype_agent, y = first_fix_old_goal_mean)) +
  geom_bar(stat="identity") + 
  facet_grid(agent_first~session_withinagent)+
  ylim(0,1)+
  ylab("Prop. dogs fixating the old goal first")+
  xlab("Agent")+
  theme(axis.text = element_text(size = 14))+  #axis text size 
  theme(axis.title = element_text(size = 16))
```

```{r}
ggsave(filename="graphs/AL_plot_NAs_excluded.png",
       plot=AL_plot_NAs_excluded,
       width=8.5,
       height=5,
       #scale=0.55,
       dpi=1200)
```


## latency of anticipatory looks by IA
New approach: use latency of anticipatory looks to old goal and old site as response variable and we include IA and Agent and their interaction as predictors. 

```{r}
first4s.data.latency.first.fix <- first4s.data %>%
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  select(
    subject,
    session_number,
    IA_fam_object ,
    condition,
    trial_type,
    trialtype_agent,
    agent_target_side_first,
    IA_FIRST_FIXATION_TIME,
    attention_getter_duration
  ) %>%
  mutate(IA_FIRST_FIXATION_TIME_corrected = IA_FIRST_FIXATION_TIME - attention_getter_duration)%>%
  select(-IA_FIRST_FIXATION_TIME, -attention_getter_duration)%>%
  pivot_wider(names_from = IA_fam_object, values_from = IA_FIRST_FIXATION_TIME_corrected)  %>%
      separate(
        agent_target_side_first,
        into = c("agent_first", "object_first", "side_first"),
        #shown to the dog in the first session
        sep = "_"
      ) %>%
      mutate(
        session_number = as.numeric(session_number),
        session_withinagent = as.numeric(ifelse(
          session_number >= 3, session_number - 2, session_number
        ))
      )

#numbers of trials in which the dogs looked at least at one of the two areas of interest (old goal or old side)
first4s.data.latency.first.fix %>%
  group_by(trialtype_agent, session_withinagent) %>%
  summarise(
    old_side_count = sum(!is.na(new_goal_old_side)),
    old_goal_count = sum(!is.na(old_goal_new_side)),
    both_count = sum(!is.na(new_goal_old_side) &
                       !is.na(old_goal_new_side)),
    either_count = sum(!is.na(new_goal_old_side) |
                         !is.na(old_goal_new_side))
  )
```
```{r}
first4s.data.latency.first.fix.long <- first4s.data.latency.first.fix %>%
  pivot_longer(cols = c(old_goal_new_side, new_goal_old_side), names_to = "IA", values_to = "latency") %>%
  filter(!is.na(latency))

table(first4s.data.latency.first.fix.long$subject, first4s.data.latency.first.fix.long$trialtype_agent)
table(first4s.data.latency.first.fix.long$subject, first4s.data.latency.first.fix.long$IA)
```


```{r}
xx.fe.re=fe.re.tab(fe.model="latency ~ IA * trialtype_agent",
re="(1|subject)", data=first4s.data.latency.first.fix.long)
xx.fe.re$summary
```

fit model (testing whether the latency to look at both AOIs is modulated by the interaction between AOI and agent)
```{r}

mm1.latency<-lme4::lmer(latency ~ IA * trialtype_agent +
            (1|subject), data=first4s.data.latency.first.fix.long, REML=FALSE)

```


```{r}
diagnostics.plot(mm1.latency, size.fac=2)

ranef.diagn.plot(mm1.latency)
```

Individual fixed effects
```{r}
summary(mm1.latency)
drop1(mm1.latency, test = "Chisq")
library(lmerTest)

mm1.latency.reml<-lmerTest::lmer(latency ~ IA * trialtype_agent +
            (1|subject), data=first4s.data.latency.first.fix.long, REML=TRUE)
summary(mm1.latency.reml)
#in the inanimate condition, dogs' latency to look at the old side is lower than in the human condition

```
```{r}
first4s.data.latency.first.fix.long.agg<- first4s.data.latency.first.fix.long %>%
  group_by(trialtype_agent, IA, subject)%>%
  summarise(mean_latency = mean(latency))
library(ggbeeswarm)

latency_AL_AOIs_plot<-ggplot(first4s.data.latency.first.fix.long.agg, aes(x = trialtype_agent, y = mean_latency, fill = IA)) +
  geom_boxplot(position = position_dodge(0.8), outlier.colour = "white") +
  #geom_quasirandom(method = "pseudorandom", width=0.2, position = position_dodge(0.8))+
  geom_jitter(position = position_dodge(0.8), alpha=.2)+
  scale_fill_manual(values=c("dodgerblue", "darkorange"), labels= c("old side","old goal"))+
  ylab("Mean latency to first fix.")+
  xlab("Agent")+
  theme_bw()
latency_AL_AOIs_plot
```
```{r}
ggsave(filename="graphs/latency_AL_AOIs.png",
  plot = latency_AL_AOIs_plot,
  width=8.5,
  height=5,
  scale=0.55,
  dpi=1200)
```
#model relevelled so that old goal is the new reference category
```{r}
first4s.data.latency.first.fix.long$IA_rel = relevel(as.factor(first4s.data.latency.first.fix.long$IA), ref = "old_goal_new_side")

levels(first4s.data.latency.first.fix.long$IA_rel)

mm1.latency.reml_rel<-lmerTest::lmer(latency ~ IA_rel * trialtype_agent +
            (1|subject), data=first4s.data.latency.first.fix.long, REML=TRUE)
summary(mm1.latency.reml_rel)
```
#model relevelled so that agent inanimate is the reference category
```{r}
first4s.data.latency.first.fix.long$trialtype_agent_rel = relevel(as.factor(first4s.data.latency.first.fix.long$trialtype_agent), ref = "inanimate")

levels(first4s.data.latency.first.fix.long$trialtype_agent_rel)

mm1.latency.reml_rel2<-lmerTest::lmer(latency ~ IA * trialtype_agent_rel +
            (1|subject), data=first4s.data.latency.first.fix.long, REML=TRUE)
summary(mm1.latency.reml_rel2)
#In the inanimate condition, dogs' latency to look at the old goal is larger than to look at the old side
```




## Proportion dwell time
```{r}

old_goal_object_data <- first4s.data %>%
  select(    subject,
    session_number,
    IA_fam_object ,
    condition,
    trial_type,
    trialtype_agent, 
    ia_object) %>%
  filter(IA_fam_object == "old_goal_new_side")%>%
  select(-IA_fam_object)%>%
  rename(old_goal_object = ia_object)


first4s.dwell.time <- first4s.data %>%
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  select(
    subject,
    session_number,
    IA_fam_object ,
    condition,
    trial_type,
    trialtype_agent,
    agent_target_side_first,
    IA_DWELL_TIME
  ) %>%
  pivot_wider(names_from = IA_fam_object, values_from = IA_DWELL_TIME)  %>%
      separate(
        agent_target_side_first,
        into = c("agent_first", "object_first", "side_first"),
        #shown to the dog in the first session
        sep = "_"
      ) %>%
      mutate(
        session_number = as.numeric(session_number),
        session_withinagent = as.numeric(ifelse(
          session_number >= 3, session_number - 2, session_number
        ))
      ) %>%
  mutate(prop_time_old_goal = old_goal_new_side / (old_goal_new_side + new_goal_old_side))%>%
  full_join(old_goal_object_data)
```

### beta model
```{r}

first4s.dwell.time$z.session_within<-as.vector(scale(first4s.dwell.time$session_withinagent, center = TRUE, scale=TRUE))
first4s.dwell.time$trialtype_agent<- as.factor(first4s.dwell.time$trialtype_agent)
first4s.dwell.time$trialtype_agent.c=as.vector(scale(as.numeric(first4s.dwell.time$trialtype_agent==levels(as.factor(first4s.dwell.time$trialtype_agent))[2]), center=TRUE, scale= FALSE))



first4s.dwell.time$prop_time_old_goal.scaled <- (first4s.dwell.time$prop_time_old_goal*(length(first4s.dwell.time$prop_time_old_goal) - 1) + 0.5)/length(first4s.dwell.time$prop_time_old_goal)#transform DV to exclude 0 and 1s


```


```{r}
library(glmmTMB)
mm1.beta<-glmmTMB(prop_time_old_goal.scaled~trialtype_agent+z.session_within+agent_first+ old_goal_object + 
         #   (1+trialtype_agent.c+z.session_within+old_goal_object|subject),
            (1|subject)+
            (0+trialtype_agent.c|subject)+
            (0+z.session_within|subject),
              #(0+old_goal_object|subject),#pruned due to warning message
            family=beta_family, data=first4s.dwell.time,  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))

overdisp.test(mm1.beta)

```
```{r}
drop1_mm1.beta<-drop1(mm1.beta, test="Chisq")%>%filter(Df!="")%>%add_row(Df = rep(NA,1),  .before = 1)
drop1_mm1.beta
```

```{r}
summary(mm1.beta)
```

model stability
```{r}

mm1.beta.stab=glmmTMB.stab(model.res=mm1.beta, para=F, data=first4s.dwell.time)

mm1.beta.stab$summary

m.stab.plot(round(mm1.beta.stab$summary[1:5, -1], 3))
#is.re=grepl(x=rownames(mm1.beta.stab$summary), pattern="@")
#m.stab.plot(mm1.beta.stab$summary[!is.re, -1])
```



```{r}
library(car)
xx=lm(prop_time_old_goal.scaled~trialtype_agent+z.session_within+agent_first+ old_goal_object, data=first4s.dwell.time)
vif(xx)
```


```{r}
mm1.beta.for.ci<-glmmTMB(prop_time_old_goal.scaled~trialtype_agent+z.session_within+agent_first+ old_goal_object + 
         #   (1+trialtype_agent.c+z.session_within+old_goal_object|subject),
            (1|subject)+
            (0+trialtype_agent.c|subject)+
            (0+z.session_within|subject),
              #(0+old_goal_object|subject),#pruned due to warning message
            family=beta_family, data=first4s.dwell.time%>%filter(!is.na(prop_time_old_goal.scaled)),  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))

boot.mm1.beta<-boot.glmmtmb(mm1.beta.for.ci, data=as.data.frame(first4s.dwell.time%>%filter(!is.na(prop_time_old_goal.scaled))),
nboots=1000, para=T, n.cores="all-1", resol=100, level=0.95 )

boot.mm1.beta$ci.estimates$fe

```



### output table

```{r}

model_table_dwell_time<- bind_cols(as.data.frame(summary(mm1.beta)$coefficients$cond),
                             boot.mm1.beta$ci.estimates$fe[1:5,],
                             drop1_mm1.beta) %>%
                             select(Estimate, SE = `Std. Error`, LowerCI = X2.5., UpperCI = X97.5.,z=`z value`, p=`Pr(>|z|)`,  Chi2 = LRT, df = Df, p_LRT = `Pr(>Chi)`) %>%
  mutate(across(.cols = c(p, p_LRT), ~ format(round(.x, 3), nsmall=3))) %>% 
  mutate(across(.cols = c(Estimate:z, Chi2), ~ format(round(.x, 2), nsmall=2))) %>% 
#  mutate(across(Chi2:p_LRT, ~replace_na(.x, "")))%>%
mutate(p_LRT=replace(p_LRT, p_LRT==0.000, "<0.001"))%>%
mutate(p=replace(p, p==0.000, "<0.001"))

write.csv(model_table_dwell_time, file = "saves/mm1_beta_prop_looking_time_IAobject_IPfirst4s.csv")
```






## old stuff
######################################################### copied from different projects ###########



```{r}
first4s.data.wider <- first4s.data %>%
  select(
    subject,
    session.num,
    condition_trial,
    Trial_Index_,
    IA_LABEL,
    IA_fam_object,
    condition,
    trial_type,
    agent_target_side_first,
    condition_first,
    video_file,
    DATA_FILE,
    IA_AREA,
    IA_DWELL_TIME,
    IA_FIXATION_COUNT
  ) %>% 
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  pivot_wider(names_from = IA_fam_object, values_from = )%>% #IA_DWELL_TIME for dwell times
  # mutate(object_congruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_lower_object, ifelse(target_location=="down",IA_upper_object, "") )), object_incongruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_upper_object, ifelse(target_location=="down",IA_lower_object, "") )))%>%
  # select(-IA_lower_object, -IA_upper_object, -IA_agent)%>%
  # pivot_longer(cols=c(object_congruent_dwell_time, object_incongruent_dwell_time), names_to = "object_congruent", values_to = "dwell_time")


#checks:
view(dfSummary(end.ip.data))

table(end.ip.data$trial_within_block, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$subject)
table(end.ip.data$session.num, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$TRIAL_TOTAL_VISITED_IA_COUNT)

```


```{r}

test.IP.end.data <- end.ip.data %>%
  group_by(subject, condition, target_location, IA_LABEL)%>%
  summarise(mean_dwell=mean(IA_DWELL_TIME))



ggplot(data = test.IP.end.data, aes(x=target_location, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  #geom_point(aes(x=first_target_loc, y=IA_DWELL_TIME, group=IA_LABEL),position_dodge(), alpha=0.5)+
  facet_grid(condition~IA_LABEL)+
  ylab("Dwell time")+xlab("Condition")+
  #ylim(0,3000)+
  theme_bw()#+
  #stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```

```{r}

test.IP.end.data.object.congruent <- end.ip.data.wider %>%
  group_by(subject, condition, object_congruent)%>%
  summarise(mean_dwell=mean(dwell_time))



ggplot(data = test.IP.end.data.object.congruent, aes(x=object_congruent, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  geom_point(alpha=0.3)+
  geom_line(aes(group=subject), alpha=0.3)+
  facet_wrap(~condition)+
  ylab("Dwell time")+xlab("Object congruency")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```


```{r}
test.IP.end.data.object.congruent.hand<-test.IP.end.data.object.congruent%>%
  filter(condition=="hand")

t.test(test.IP.end.data.object.congruent.hand$mean_dwell[test.IP.end.data.object.congruent.hand$object_congruent=="object_congruent_dwell_time"], test.IP.end.data.object.congruent.hand$mean_dwell[test.IP.end.data.object.congruent.hand$object_congruent=="object_incongruent_dwell_time"], paired=TRUE)
```




End IP (longer)

```{r}
end.ip.long.orig.data <- read_delim("data/dog_woodward_paradigm_IA_report_video_IP_end_long.txt", na=".", delim="\t")

demo.data <- read_csv("data/Woodward_Kano_ counterbalancing.csv")%>%
  mutate(EDF1=fct_recode(as.factor(EDF1), "Georgia_1"="Georgia1"))%>%
  separate(EDF1, c("subject", "num"), sep="_")


end.ip.long.data <- end.ip.long.orig.data %>%
  select(Session_Name_, IP_LABEL, IA_LABEL,IA_ID, Trial_Index_, condition, target, target_location, session, trial_within_block, video_file,  video_latency, DATA_FILE,  IA_AREA, IA_AVERAGE_FIX_PUPIL_SIZE,  IA_DWELL_TIME, "IA_DWELL_TIME_%", IA_FIRST_FIXATION_TIME, IA_FIXATION_COUNT, IA_MAX_FIX_PUPIL_SIZE, INTEREST_AREA_FIXATION_SEQUENCE, TRIAL_DWELL_TIME, TRIAL_FIXATION_COUNT, TRIAL_IA_COUNT, TRIAL_TOTAL_VISITED_IA_COUNT )%>%
 #  filter(!(DATA_FILE=="Ace_2.edf" & Trial_Index_!=4))%>%
#  mutate(trial=ifelse(DATA_FILE=="Ace_2.edf" & Trial_Index_==4, 5, Trial_Index_))%>% #recoding of trial number to account for a dog that did not complete his first session, session was repeated with this dog
  mutate(Session_Name_=fct_recode(Session_Name_, Georgia_1="Georgia1",Georgia_2="Georgia2", Georgia_3="Georgia3",Georgia_4="Georgia4" ))%>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial=paste(condition, "_", trial_within_block))%>%
  full_join(demo.data)
  


end.ip.long.data.wider<-end.ip.long.data%>%
  select(subject, session.num,condition_trial, IP_LABEL, IA_LABEL, condition, target, target_location, session, trial_within_block, video_file,  video_latency, DATA_FILE,  IA_DWELL_TIME)%>%
  pivot_wider(names_from = IA_LABEL, values_from=IA_DWELL_TIME)%>%
  mutate(object_congruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_lower_object, ifelse(target_location=="down",IA_upper_object, "") )), object_incongruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_upper_object, ifelse(target_location=="down",IA_lower_object, "") )))%>%
  select(-IA_lower_object, -IA_upper_object, -IA_agent)%>%
  pivot_longer(cols=c(object_congruent_dwell_time, object_incongruent_dwell_time), names_to = "object_congruent", values_to = "dwell_time")


#checks:
view(dfSummary(end.ip.data))

table(end.ip.data$trial_within_block, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$subject)
table(end.ip.data$session.num, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$TRIAL_TOTAL_VISITED_IA_COUNT)

```


```{r}

test.IP.end.long.data <- end.ip.long.data %>%
  group_by(subject, condition, target_location, IA_LABEL)%>%
  summarise(mean_dwell=mean(IA_DWELL_TIME))



ggplot(data = test.IP.end.long.data, aes(x=target_location, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  #geom_point(aes(x=first_target_loc, y=IA_DWELL_TIME, group=IA_LABEL),position_dodge(), alpha=0.5)+
  facet_grid(condition~IA_LABEL)+
  ylab("Dwell time")+xlab("Condition")+
  #ylim(0,3000)+
  theme_bw()#+
  #stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```

```{r}

test.IP.end.long.data.object.congruent <- end.ip.long.data.wider %>%
  group_by(subject, condition, object_congruent)%>%
  summarise(mean_dwell=mean(dwell_time))



ggplot(data = test.IP.end.long.data.object.congruent, aes(x=object_congruent, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  geom_point(alpha=0.3)+
  geom_line(aes(group=subject), alpha=0.3)+
  facet_wrap(~condition)+
  ylab("Dwell time")+xlab("Object congruency")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```

Trial report

```{r}

trialreport.end.ip.orig.data <- read_delim("data/dog_woodward_paradigm_trial_report_video_IP_end.txt", na=".", delim="\t")



```







-repear IA analysis with larger areas of interest
- setup a model for the dwell time (proportion of correct AOI)
- first look
- pupil size







############################################################

```{r}
#first test trial
test.IP.end.IA.endpos.trial1 <- end.ip.data %>%
  filter(condition!="hab", IA_LABEL=="IA_end_position", trial==4)


t.test(test.IP.end.IA.endpos.trial1$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="exp"], test.IP.end.IA.endpos.trial1$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="con"], paired=TRUE)

#second test trial
test.IP.end.IA.endpos.trial2 <- end.ip.data %>%
  filter(condition!="hab", IA_LABEL=="IA_end_position", trial==5)


t.test(test.IP.end.IA.endpos.trial2$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="exp"], test.IP.end.IA.endpos.trial2$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="con"], paired=TRUE)


```


Onscreen Dwell time (while videos is playing)

```{r}
video_playing_data<-read.delim("data/ET_VoE_occlusion2screens_IA_report_whole_screen_video_playing_IP.txt",na=".", sep = "\t")%>%
  select(Session_Name_, IP_LABEL, IA_LABEL,IA_ID, Trial_Index_, condition, phase, session,  video_file, video_id, video_latency, DATA_FILE,  IA_AREA, IA_AVERAGE_FIX_PUPIL_SIZE,  IA_DWELL_TIME, "IA_DWELL_TIME", IA_FIXATION_COUNT, IA_MAX_FIX_PUPIL_SIZE, INTEREST_AREA_FIXATION_SEQUENCE, TRIAL_DWELL_TIME, TRIAL_FIXATION_COUNT, TRIAL_IA_COUNT, TRIAL_TOTAL_VISITED_IA_COUNT )%>%
   filter(!(DATA_FILE=="Ace_2.edf" & Trial_Index_!=4))%>%
  mutate(trial=ifelse(DATA_FILE=="Ace_2.edf" & Trial_Index_==4, 5, Trial_Index_))%>% #recoding of trial number to account for a dog that did not complete his first session, session was repeated with this dog
  mutate(Session_Name_=fct_recode(Session_Name_, Georgia_1="Georgia1",Georgia_2="Georgia2", Cheynna_1="Cheynna1", Cheynna_2="Cheynna2" ))%>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial=paste(condition, "_", trial))%>%
  full_join(demo.data)
 # rename(trial="Trial_Index_")
  

#checks:
view(dfSummary(video_playing_data))


on_screen_data<-video_playing_data%>%
  filter(IA_LABEL=="screen_aoi")%>%
  mutate(prop_looking=IA_DWELL_TIME/5000)


on_screen_data%>%group_by(condition)%>%
  summarise(mean=mean(prop_looking, na.rm = TRUE), median=median(prop_looking, na.rm = TRUE), se=sd(prop_looking, na.rm = TRUE)/sqrt(length(prop_looking)), min=min(prop_looking, na.rm = TRUE), max=max(prop_looking, na.rm = TRUE))

```

onscreen dwell-time: comparison between first and last familiarisation trial.
```{r}
library(broom)
on_screen_data%>%filter(phase=="hab1" | phase=="hab3")%>%
  group_by(subject, phase)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ phase, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```


onscreen dwell-time: comparison between test trials
* first test trial
```{r}
on_screen_data%>%filter(condition!="hab", trial==4)%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ condition, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```

* second test trial
```{r}
on_screen_data%>%filter(condition!="hab", trial==5)%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ condition, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```

Plot onscreen dwell time in test trials
```{r}
xx<-on_screen_data%>%filter(condition!="hab")%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking))%>%
  ungroup()

ggplot(data=xx, aes(x=condition, y=mean))+
  geom_boxplot()
```