---
title: Woodward 2022: Dwell time analysis
author: Christoph Voelter
date: 
output: html_document
---
notes: 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#rm(list=ls())
#remove(list=c("first4s.data_short"))
library(tidyverse)
library(summarytools)
library(lme4)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
load("workspaces/AL_test_first4s.RData")
```

## Original IAs
### End IP
```{r}
demo.data <- read_csv("data/Woodward_2022_counterbalancing.csv") %>%
  mutate(
    EDF1 = fct_recode(
      as.factor(EDF1),
      "Georgia_1" = "Georgia1",
    )
  ) %>%
  separate(EDF1, c("subject", "num"), sep = "_")
```

```{r}
first4s.data <-
  read_delim(
    "data/Woodward2022_IA_Report_IP_test_first_4_sec.txt",
    na = ".",
    delim = "\t"
  ) %>%
  dplyr::select(
    Session_Name_,
    IP_LABEL,
    IA_LABEL,
    IA_ID,
    Trial_Index_,
    condition,
    dog_session,
    sub_id,
    trial_number,
    trial_type,
    session_number,
    video_file,
    DATA_FILE,
    IA_AREA,
    IA_AVERAGE_FIX_PUPIL_SIZE,
    IA_DWELL_TIME,
    "IA_DWELL_TIME_%",
    IA_FIXATION_COUNT,
    "IA_FIXATION_%",
    IA_FIRST_FIXATION_INDEX, #Ordinal sequence of the first fixation that       #was within the current interest area.
    IA_FIRST_FIXATION_VISITED_IA_COUNT, #number of different interest areas
    #visited so far before the first fixation is made to the
    #current interest area.
    IA_MAX_FIX_PUPIL_SIZE,
    IA_FIRST_FIXATION_TIME,
    INTEREST_AREA_FIXATION_SEQUENCE,
    TRIAL_DWELL_TIME,
    TRIAL_FIXATION_COUNT,
    TRIAL_IA_COUNT,
    TRIAL_TOTAL_VISITED_IA_COUNT
  ) %>%
  filter(DATA_FILE != "Melody_1.edf", DATA_FILE != "Mia_5.edf") %>%
  filter(DATA_FILE != "Timo_5.edf", DATA_FILE != "Joker_5.edf") %>% # backup recordings in case the first sessions are below 70% onscreen looking
  #filter(!(DATA_FILE == "Amy_4.edf" & Trial_Index_ != 1)) %>%
  #mutate(session.num = as.numeric(session.num)) %>%
  #mutate(session.num.corrected = ifelse(DATA_FILE == "Amy_4.edf" | DATA_FILE == "Amy_5.edf" | DATA_FILE == "George_5.edf" | DATA_FILE == "Melody_2.edf" | DATA_FILE == #"Melody_3.edf" | DATA_FILE == "Melody_4.edf", session.num-1,
#                                        ifelse(DATA_FILE == "Melody_7.edf" | DATA_FILE == "Mia_6.edf", session.num-3, session.num))) %>% #recoding of session number to account for dogs with aborted sessions
  mutate(
    Session_Name_ = fct_recode(
      Session_Name_,
      Georgia_1 = "Georgia1",
      Georgia_2 = "Georgia2",
      Georgia_3 = "Georgia3",
      Georgia_4 = "Georgia4"
    )
  ) %>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial = paste(condition, "_", Trial_Index_)) %>%
  full_join(demo.data) %>%
  mutate(IA_LABEL2 = IA_LABEL, trial_type2 = trial_type) %>%
  separate(
    IA_LABEL2,
    into = c("ia_object", "ia_side", "ia_label_del", "ia_label_del2"),
    sep = "_"
  ) %>%
  separate(
    trial_type2,
    into = c("trialtype_agent", "trialtype_object", "trialtype_side"),
    sep = "_"
  ) %>%
  filter(ia_object == "elephant" | ia_object == "ball") %>%
  mutate(
    IA_fam_object = ifelse(
      condition == "new_goal" &
        ia_object == trialtype_object,
      "new_goal_old_side",
      ifelse(
        condition == "new_goal" &
          ia_object != trialtype_object,
        "old_goal_new_side",
        ifelse(
          condition == "new_side" &
            ia_object != trialtype_object,
          "new_goal_old_side",
          ifelse(
            condition == "new_side" &
              ia_object == trialtype_object,
            "old_goal_new_side",
            ""
          )
        )
      )
    )
  ) 

first4s.data %>% select(subject, condition_trial, trial_type, IA_LABEL, IA_fam_object) 
#IA_fam_object=what the AOI looked at contains
first4s.data %>% select(subject, condition, trial_type, IA_LABEL, IA_fam_object,IA_FIRST_FIXATION_INDEX,IA_FIRST_FIXATION_VISITED_IA_COUNT) 


levels(as.factor(first4s.data$DATA_FILE))
levels(as.factor(first4s.data$condition))
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$Trial_Index_)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$Trial_Index_)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$session.num)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$session_number)
```

```{r}

first4s.data.first.fix <- first4s.data %>%
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  select(subject, session_number, IA_fam_object , condition, ,trial_type, trialtype_agent, agent_target_side_first, IA_FIRST_FIXATION_TIME)%>%
  pivot_wider(names_from = IA_fam_object, values_from=IA_FIRST_FIXATION_TIME)%>%
  mutate(old_goal_new_side = replace_na(old_goal_new_side, 99999999), new_goal_old_side = replace_na(new_goal_old_side, 99999999))%>%
  mutate(first_fix_old_goal=as.numeric(ifelse(old_goal_new_side<new_goal_old_side, 1, 0)),
         first_fix_old_goal_withoutNA = as.numeric(ifelse(old_goal_new_side==new_goal_old_side, NA,
                                                          ifelse(old_goal_new_side<new_goal_old_side, 1, 0)))) %>%
    separate(
    agent_target_side_first,
    into = c("agent_first", "object_first", "side_first"), #shown to the dog in the first session
    sep = "_"
  ) %>%
  mutate(session_number = as.numeric(session_number), session_withinagent = as.numeric(ifelse(session_number >=3, session_number-2, session_number)))


```


## first fixation (binary model and no fixations are counted as 0)
```{r}

first4s.data.first.fix$z.session <-
  as.vector(scale(
    as.numeric(first4s.data.first.fix$session_number),
    center = TRUE,
    scale = TRUE
  ))

first4s.data.first.fix$trialtype_agent <-
  as.factor(first4s.data.first.fix$trialtype_agent)
levels(first4s.data.first.fix$trialtype_agent)

first4s.data.first.fix$trialtype_agent.c = as.vector(scale(
  as.numeric(
    first4s.data.first.fix$trialtype_agent == levels(as.factor(first4s.data.first.fix$trialtype_agent))[2]
  ),
  center = TRUE,
  scale = FALSE
))

first4s.data.first.fix$agent_first.c = as.vector(scale(
  as.numeric(first4s.data.first.fix$agent_first == levels(
    as.factor(first4s.data.first.fix$agent_first)
  )[2]),
  center = TRUE,
  scale = FALSE
))
levels(as.factor(first4s.data.first.fix$z.session))
hist(first4s.data.first.fix$first_fix_old_goal)
```

fit model
```{r}
mm1_first_fix=glmer(first_fix_old_goal ~ trialtype_agent+z.session+agent_first +
            (1+trialtype_agent.c + z.session|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#prune random slope structure due to convergence warning (i.e, exclude correlation between random icpt and slopes, remove random slope of session number)
mm1_first_fix2=glmer(first_fix_old_goal ~ trialtype_agent+z.session+agent_first +
            (1|subject)+(0+trialtype_agent.c |subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(mm1_first_fix2)
```


```{r}

mm1_first_fix2_drop1 <- drop1(mm1_first_fix2, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1)
```


Check for colinearity
```{r}
library(car)
xx=lm(first_fix_old_goal ~ trialtype_agent+z.session+agent_first, data=first4s.data.first.fix)
vif(xx)
```
#### Confidence intervals of the estimates
```{r}
mm1_first_fix2.ci=boot.glmm.pred(model.res=mm1_first_fix2, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)
mm1_first_fix2.ci$ci.estimates

```

#### model stability
```{r}
m.stab.mm1_first_fix2 <- glmm.model.stab(model.res = mm1_first_fix2)
m.stab.mm1_first_fix2$detailed$warnings
m.stab.mm1_first_fix2_output <- as.data.frame(round(m.stab.mm1_first_fix2$summary[, -1], 3))
m.stab.mm1_first_fix2_output 
png("graphs/mm1_first_fix2_stability_plot.png")
m.stab.plot(round(m.stab.mm1_first_fix2$summary[, -1], 3))
dev.off()
m.stab.plot(round(m.stab.mm1_first_fix2$summary[, -1], 3))
# The model appeared to be stable with respect to the fixed effects (see full_model_stability_plot).

```


#### output table


```{r}
mm1_first_fix2_output_table <-
  bind_cols(as.data.frame(summary(mm1_first_fix2)$coeff),
            mm1_first_fix2_drop1,
            mm1_first_fix2.ci$ci.estimates,
           m.stab.mm1_first_fix2_output[1:4,]) %>%
  select(
    Estimate,
    SE = `Std. Error`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`,
    min_stab = min,
    max_stab = max
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  mutate(across(.cols = c(z_wald:max_stab), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_first_fix2_output_table, file = "saves/mm1_first_fix2_output_table.csv")
```

## Plot showing the object looks in the first 4 s of the test trials (including dogs that did not look at any of the objects in a given trial)
```{r}
first4s.data.first.fix.nolooked.included.agg<-first4s.data.first.fix %>%
  group_by(trialtype_agent, session_withinagent) %>%
  summarise(first_fix_old_goal_mean = mean(first_fix_old_goal, na.rm=TRUE), 
            first_fix_old_goal_sum = sum(first_fix_old_goal==1, na.rm=TRUE), 
            first_fix_old_goal_count = sum(!is.na(first_fix_old_goal)))

ggplot(first4s.data.first.fix.nolooked.included.agg, aes(x = trialtype_agent, y = first_fix_old_goal_mean)) +
  geom_bar(stat="identity") + 
  facet_wrap(~session_withinagent)+
  ylim(0,1)
```


## first fixation (binary model and no fixations are treated as NA)
```{r}
hist(first4s.data.first.fix$first_fix_old_goal_withoutNA)
```

fit model
```{r}
mm1_first_fix_withoutNA=glmer(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session+agent_first +
            (1+trialtype_agent.c + z.session|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
#prune random slope structure due to convergence warning
mm1_first_fix_withoutNA2=glmer(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session+agent_first +
            (1|subject)+(0+trialtype_agent.c + z.session|subject),#removed the correlation between random intercept and random slope
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
#model has unusually large predictor estimates and variance of the random slopes
summary(mm1_first_fix_withoutNA2)
mm1_first_fix_withoutNA2_drop1<-drop1(mm1_first_fix_withoutNA2, test="Chisq")
```

pruned model (random slope of z.session removed)
```{r}
mm1_first_fix_withoutNA2_pruned=glmer(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),#removed the correlation between random intercept and random slope and the random slope of session
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
# estimates and random slope variance look better (values are smaller)
#null model because drop 1 does not converge
mm1_first_fix_withoutNA2_pruned_null=glmer(first_fix_old_goal_withoutNA ~ z.session+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),#removed the correlation between random intercept and random slope
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
anova(mm1_first_fix_withoutNA2_pruned, mm1_first_fix_withoutNA2_pruned_null, test = "Chisq")
#--> agent not significant 
summary(mm1_first_fix_withoutNA2_pruned)
drop1(mm1_first_fix_withoutNA2_pruned, test = "Chisq") #does not converge
```



Check for collinearity - model without NAs
```{r}
library(car)
xx=lm(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session+agent_first, data=first4s.data.first.fix)
vif(xx)
```
#### Confidence intervals of the estimates
```{r}
mm1_first_fix2_withoutNAs.ci=boot.glmm.pred(model.res=mm1_first_fix_withoutNA2, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)
mm1_first_fix2_withoutNAs.ci$ci.estimates

```

#### model stability
```{r}
m.stab.mm1_first_fix2_withoutNAs <- glmm.model.stab(model.res = mm1_first_fix_withoutNA2)
m.stab.mm1_first_fix2_withoutNAs$detailed$warnings #models failed to converge
m.stab.mm1_first_fix2_withoutNAs_output <- as.data.frame(round(m.stab.mm1_first_fix2_withoutNAs$summary[, -1], 3))
m.stab.mm1_first_fix2_withoutNAs_output 
png("graphs/mm1_first_fix2_withoutNAs_stability_plot.png")
m.stab.plot(round(m.stab.mm1_first_fix2_withoutNAs$summary[, -1], 3))
#dev.off()
# The estimates do not look reasonable
```


#### output table


```{r}
mm1_first_fix2_withoutNAs_output_table <-
  bind_cols(as.data.frame(summary(mm1_first_fix_withoutNA2)$coeff),
            mm1_first_fix_withoutNA2_drop1,
            mm1_first_fix2_withoutNAs.ci$ci.estimates,
            m.stab.mm1_first_fix2_withoutNAs_output[1:4,]) %>%
  select(
    Estimate,
    SE = `Std. Error`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`,
    min_stab = min,
    max_stab = max
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  mutate(across(.cols = c(z_wald:max_stab), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_first_fix2_withoutNAs_output_table, file = "saves/mm1_first_fix2_withoutNAs_output_table.csv")
```

## Plot showing the object looks in the first 4 s of the test trials (without dogs that did not look not at any of the objects in a given trial)
```{r}
first4s.data.first.fix.agg<-first4s.data.first.fix %>%
  group_by(trialtype_agent, agent_first, session_withinagent) %>%
  summarise(first_fix_old_goal_mean = mean(first_fix_old_goal_withoutNA, na.rm=TRUE), #number of dogs that anticipated the old goal object
            first_fix_old_goal_sum = sum(first_fix_old_goal_withoutNA==1, na.rm=TRUE), #number of dogs that looked at at least one of the objects
            first_fix_old_goal_count = sum(!is.na(first_fix_old_goal_withoutNA)))
```


```{r}
first4s.data.first.fix.agg2<-first4s.data.first.fix %>%
  group_by(trialtype_agent) %>%
  summarise(first_fix_old_goal_mean = mean(first_fix_old_goal_withoutNA, na.rm=TRUE), 
            first_fix_old_goal_sum = sum(first_fix_old_goal_withoutNA==1, na.rm=TRUE), #number of dogs that anticipated the old goal object
            first_fix_old_side_sum = sum(first_fix_old_goal_withoutNA==0, na.rm=TRUE), #number of dogs that anticipated the old goal object
            first_fix_old_goal_count = sum(!is.na(first_fix_old_goal_withoutNA)))#number of dogs that looked at at least one of the objects

#binomial test for
#human agent
binom.test(x = first4s.data.first.fix.agg2$first_fix_old_goal_sum[1] , first4s.data.first.fix.agg2$first_fix_old_goal_count[1] )
#and for inanimate agent
binom.test(x = first4s.data.first.fix.agg2$first_fix_old_goal_sum[2] , first4s.data.first.fix.agg2$first_fix_old_goal_count[2] )
```

```{r}

ggplot(first4s.data.first.fix.agg, aes(x = trialtype_agent, y = first_fix_old_goal_mean)) +
  geom_bar(stat="identity") + 
  facet_grid(agent_first~session_withinagent)+
  ylim(0,1)
```

## intercept only model ( to test whether the dogs have a preference for the new goal in the human condition)

```{r}
mm1_first_fix_intercept_only=glmer(first_fix_old_goal_withoutNA ~ +z.session+agent_first.c +
            (1|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mm1_first_fix_intercept_only)
```




######################################################### copied from different projects ###########



```{r}
first4s.data.wider <- first4s.data %>%
  select(
    subject,
    session.num,
    condition_trial,
    Trial_Index_,
    IA_LABEL,
    IA_fam_object,
    condition,
    trial_type,
    agent_target_side_first,
    condition_first,
    video_file,
    DATA_FILE,
    IA_AREA,
    IA_DWELL_TIME,
    IA_FIXATION_COUNT
  ) %>% 
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  pivot_wider(names_from = IA_fam_object, values_from = )%>% #IA_DWELL_TIME for dwell times
  # mutate(object_congruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_lower_object, ifelse(target_location=="down",IA_upper_object, "") )), object_incongruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_upper_object, ifelse(target_location=="down",IA_lower_object, "") )))%>%
  # select(-IA_lower_object, -IA_upper_object, -IA_agent)%>%
  # pivot_longer(cols=c(object_congruent_dwell_time, object_incongruent_dwell_time), names_to = "object_congruent", values_to = "dwell_time")


#checks:
view(dfSummary(end.ip.data))

table(end.ip.data$trial_within_block, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$subject)
table(end.ip.data$session.num, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$TRIAL_TOTAL_VISITED_IA_COUNT)

```


```{r}

test.IP.end.data <- end.ip.data %>%
  group_by(subject, condition, target_location, IA_LABEL)%>%
  summarise(mean_dwell=mean(IA_DWELL_TIME))



ggplot(data = test.IP.end.data, aes(x=target_location, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  #geom_point(aes(x=first_target_loc, y=IA_DWELL_TIME, group=IA_LABEL),position_dodge(), alpha=0.5)+
  facet_grid(condition~IA_LABEL)+
  ylab("Dwell time")+xlab("Condition")+
  #ylim(0,3000)+
  theme_bw()#+
  #stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```

```{r}

test.IP.end.data.object.congruent <- end.ip.data.wider %>%
  group_by(subject, condition, object_congruent)%>%
  summarise(mean_dwell=mean(dwell_time))



ggplot(data = test.IP.end.data.object.congruent, aes(x=object_congruent, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  geom_point(alpha=0.3)+
  geom_line(aes(group=subject), alpha=0.3)+
  facet_wrap(~condition)+
  ylab("Dwell time")+xlab("Object congruency")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```


```{r}
test.IP.end.data.object.congruent.hand<-test.IP.end.data.object.congruent%>%
  filter(condition=="hand")

t.test(test.IP.end.data.object.congruent.hand$mean_dwell[test.IP.end.data.object.congruent.hand$object_congruent=="object_congruent_dwell_time"], test.IP.end.data.object.congruent.hand$mean_dwell[test.IP.end.data.object.congruent.hand$object_congruent=="object_incongruent_dwell_time"], paired=TRUE)
```




### End IP (longer)

```{r}
end.ip.long.orig.data <- read_delim("data/dog_woodward_paradigm_IA_report_video_IP_end_long.txt", na=".", delim="\t")

demo.data <- read_csv("data/Woodward_Kano_ counterbalancing.csv")%>%
  mutate(EDF1=fct_recode(as.factor(EDF1), "Georgia_1"="Georgia1"))%>%
  separate(EDF1, c("subject", "num"), sep="_")


end.ip.long.data <- end.ip.long.orig.data %>%
  select(Session_Name_, IP_LABEL, IA_LABEL,IA_ID, Trial_Index_, condition, target, target_location, session, trial_within_block, video_file,  video_latency, DATA_FILE,  IA_AREA, IA_AVERAGE_FIX_PUPIL_SIZE,  IA_DWELL_TIME, "IA_DWELL_TIME_%", IA_FIRST_FIXATION_TIME, IA_FIXATION_COUNT, IA_MAX_FIX_PUPIL_SIZE, INTEREST_AREA_FIXATION_SEQUENCE, TRIAL_DWELL_TIME, TRIAL_FIXATION_COUNT, TRIAL_IA_COUNT, TRIAL_TOTAL_VISITED_IA_COUNT )%>%
 #  filter(!(DATA_FILE=="Ace_2.edf" & Trial_Index_!=4))%>%
#  mutate(trial=ifelse(DATA_FILE=="Ace_2.edf" & Trial_Index_==4, 5, Trial_Index_))%>% #recoding of trial number to account for a dog that did not complete his first session, session was repeated with this dog
  mutate(Session_Name_=fct_recode(Session_Name_, Georgia_1="Georgia1",Georgia_2="Georgia2", Georgia_3="Georgia3",Georgia_4="Georgia4" ))%>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial=paste(condition, "_", trial_within_block))%>%
  full_join(demo.data)
  


end.ip.long.data.wider<-end.ip.long.data%>%
  select(subject, session.num,condition_trial, IP_LABEL, IA_LABEL, condition, target, target_location, session, trial_within_block, video_file,  video_latency, DATA_FILE,  IA_DWELL_TIME)%>%
  pivot_wider(names_from = IA_LABEL, values_from=IA_DWELL_TIME)%>%
  mutate(object_congruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_lower_object, ifelse(target_location=="down",IA_upper_object, "") )), object_incongruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_upper_object, ifelse(target_location=="down",IA_lower_object, "") )))%>%
  select(-IA_lower_object, -IA_upper_object, -IA_agent)%>%
  pivot_longer(cols=c(object_congruent_dwell_time, object_incongruent_dwell_time), names_to = "object_congruent", values_to = "dwell_time")


#checks:
view(dfSummary(end.ip.data))

table(end.ip.data$trial_within_block, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$subject)
table(end.ip.data$session.num, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$TRIAL_TOTAL_VISITED_IA_COUNT)

```


```{r}

test.IP.end.long.data <- end.ip.long.data %>%
  group_by(subject, condition, target_location, IA_LABEL)%>%
  summarise(mean_dwell=mean(IA_DWELL_TIME))



ggplot(data = test.IP.end.long.data, aes(x=target_location, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  #geom_point(aes(x=first_target_loc, y=IA_DWELL_TIME, group=IA_LABEL),position_dodge(), alpha=0.5)+
  facet_grid(condition~IA_LABEL)+
  ylab("Dwell time")+xlab("Condition")+
  #ylim(0,3000)+
  theme_bw()#+
  #stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```

```{r}

test.IP.end.long.data.object.congruent <- end.ip.long.data.wider %>%
  group_by(subject, condition, object_congruent)%>%
  summarise(mean_dwell=mean(dwell_time))



ggplot(data = test.IP.end.long.data.object.congruent, aes(x=object_congruent, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  geom_point(alpha=0.3)+
  geom_line(aes(group=subject), alpha=0.3)+
  facet_wrap(~condition)+
  ylab("Dwell time")+xlab("Object congruency")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```

### Trial report

```{r}

trialreport.end.ip.orig.data <- read_delim("data/dog_woodward_paradigm_trial_report_video_IP_end.txt", na=".", delim="\t")



```







-repear IA analysis with larger areas of interest
- setup a model for the dwell time (proportion of correct AOI)
- first look
- pupil size







############################################################

```{r}
#first test trial
test.IP.end.IA.endpos.trial1 <- end.ip.data %>%
  filter(condition!="hab", IA_LABEL=="IA_end_position", trial==4)


t.test(test.IP.end.IA.endpos.trial1$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="exp"], test.IP.end.IA.endpos.trial1$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="con"], paired=TRUE)

#second test trial
test.IP.end.IA.endpos.trial2 <- end.ip.data %>%
  filter(condition!="hab", IA_LABEL=="IA_end_position", trial==5)


t.test(test.IP.end.IA.endpos.trial2$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="exp"], test.IP.end.IA.endpos.trial2$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="con"], paired=TRUE)


```


### Onscreen Dwell time (while videos is playing)

```{r}
video_playing_data<-read.delim("data/ET_VoE_occlusion2screens_IA_report_whole_screen_video_playing_IP.txt",na=".", sep = "\t")%>%
  select(Session_Name_, IP_LABEL, IA_LABEL,IA_ID, Trial_Index_, condition, phase, session,  video_file, video_id, video_latency, DATA_FILE,  IA_AREA, IA_AVERAGE_FIX_PUPIL_SIZE,  IA_DWELL_TIME, "IA_DWELL_TIME", IA_FIXATION_COUNT, IA_MAX_FIX_PUPIL_SIZE, INTEREST_AREA_FIXATION_SEQUENCE, TRIAL_DWELL_TIME, TRIAL_FIXATION_COUNT, TRIAL_IA_COUNT, TRIAL_TOTAL_VISITED_IA_COUNT )%>%
   filter(!(DATA_FILE=="Ace_2.edf" & Trial_Index_!=4))%>%
  mutate(trial=ifelse(DATA_FILE=="Ace_2.edf" & Trial_Index_==4, 5, Trial_Index_))%>% #recoding of trial number to account for a dog that did not complete his first session, session was repeated with this dog
  mutate(Session_Name_=fct_recode(Session_Name_, Georgia_1="Georgia1",Georgia_2="Georgia2", Cheynna_1="Cheynna1", Cheynna_2="Cheynna2" ))%>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial=paste(condition, "_", trial))%>%
  full_join(demo.data)
 # rename(trial="Trial_Index_")
  

#checks:
view(dfSummary(video_playing_data))


on_screen_data<-video_playing_data%>%
  filter(IA_LABEL=="screen_aoi")%>%
  mutate(prop_looking=IA_DWELL_TIME/5000)


on_screen_data%>%group_by(condition)%>%
  summarise(mean=mean(prop_looking, na.rm = TRUE), median=median(prop_looking, na.rm = TRUE), se=sd(prop_looking, na.rm = TRUE)/sqrt(length(prop_looking)), min=min(prop_looking, na.rm = TRUE), max=max(prop_looking, na.rm = TRUE))

```

onscreen dwell-time: comparison between first and last familiarisation trial.
```{r}
library(broom)
on_screen_data%>%filter(phase=="hab1" | phase=="hab3")%>%
  group_by(subject, phase)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ phase, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```


onscreen dwell-time: comparison between test trials
* first test trial
```{r}
on_screen_data%>%filter(condition!="hab", trial==4)%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ condition, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```

* second test trial
```{r}
on_screen_data%>%filter(condition!="hab", trial==5)%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ condition, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```

Plot onscreen dwell time in test trials
```{r}
xx<-on_screen_data%>%filter(condition!="hab")%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking))%>%
  ungroup()

ggplot(data=xx, aes(x=condition, y=mean))+
  geom_boxplot()
```