---
title: Woodward 2022: Dwell time analysis
author: Christoph Voelter
date: 
output: html_document
---
notes: 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#rm(list=ls())
#remove(list=c("first4s.data_short"))
library(tidyverse)
library(summarytools)
#library(exactRankTests)
#library(ggpubr)

```

## Original IAs
### End IP
```{r}
demo.data <- read_csv("data/Woodward_2022_counterbalancing.csv") %>%
  mutate(
    EDF1 = fct_recode(
      as.factor(EDF1),
      "Georgia_1" = "Georgia1",
    )
  ) %>%
  separate(EDF1, c("subject", "num"), sep = "_")
```

```{r}
first4s.data <-
  read_delim(
    "data/Woodward2022_IA_Report_IP_test_first_4_sec.txt",
    na = ".",
    delim = "\t"
  ) %>%
  dplyr::select(
    Session_Name_,
    IP_LABEL,
    IA_LABEL,
    IA_ID,
    Trial_Index_,
    condition,
    dog_session,
    sub_id,
    trial_number,
    trial_type,
    session_number,
    video_file,
    DATA_FILE,
    IA_AREA,
    IA_AVERAGE_FIX_PUPIL_SIZE,
    IA_DWELL_TIME,
    "IA_DWELL_TIME_%",
    IA_FIXATION_COUNT,
    "IA_FIXATION_%",
    IA_FIRST_FIXATION_INDEX, #Ordinal sequence of the first fixation that       #was within the current interest area.
    IA_FIRST_FIXATION_VISITED_IA_COUNT, #number of different interest areas
    #visited so far before the first fixation is made to the
    #current interest area.
    IA_MAX_FIX_PUPIL_SIZE,
    INTEREST_AREA_FIXATION_SEQUENCE,
    TRIAL_DWELL_TIME,
    TRIAL_FIXATION_COUNT,
    TRIAL_IA_COUNT,
    TRIAL_TOTAL_VISITED_IA_COUNT
  ) %>%
  filter(DATA_FILE != "Melody_1.edf", DATA_FILE != "Mia_5.edf") %>%
  filter(DATA_FILE != "Timo_5.edf", DATA_FILE != "Joker_5.edf") %>% # backup recordings in case the first sessions are below 70% onscreen looking
  #filter(!(DATA_FILE == "Amy_4.edf" & Trial_Index_ != 1)) %>%
  # mutate(trial = ifelse(DATA_FILE == "Ace_.edf" &
  #                         Trial_Index_ == 4, 5, Trial_Index_))#recoding of trial number to account for a dog that did not complete his first session, session was repeated with this dog
  mutate(
    Session_Name_ = fct_recode(
      Session_Name_,
      Georgia_1 = "Georgia1",
      Georgia_2 = "Georgia2",
      Georgia_3 = "Georgia3",
      Georgia_4 = "Georgia4"
    )
  ) %>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial = paste(condition, "_", Trial_Index_)) %>%
  full_join(demo.data) %>%
  mutate(IA_LABEL2 = IA_LABEL, trial_type2 = trial_type) %>%
  separate(
    IA_LABEL2,
    into = c("ia_object", "ia_side", "ia_label_del", "ia_label_del2"),
    sep = "_"
  ) %>%
  separate(
    trial_type2,
    into = c("trialtype_agent", "trialtype_object", "trialtype_side"),
    sep = "_"
  ) %>%
  mutate(
    IA_fam_object = ifelse(
      condition == "new_goal" &
        ia_object == trialtype_object,
      "new_goal_old_side",
      ifelse(
        condition == "new_goal" &
          ia_object != trialtype_object,
        "old_goal_new_side",
        ifelse(
          condition == "new_side" &
            ia_object != trialtype_object,
          "new_goal_old_side",
          ifelse(
            condition == "new_side" &
              ia_object == trialtype_object,
            "old_goal_new_side",
            ""
          )
        )
      )
    )
  ) %>%
  filter(ia_object == "elephant" | ia_object == "ball")

first4s.data %>% select(subject, condition_trial, trial_type, IA_LABEL, IA_fam_object) 
#IA_fam_object=what the AOI looked at contains
first4s.data %>% select(subject, condition, trial_type, IA_LABEL, IA_fam_object,IA_FIRST_FIXATION_INDEX,IA_FIRST_FIXATION_VISITED_IA_COUNT) 


levels(as.factor(first4s.data$DATA_FILE))
levels(as.factor(first4s.data$condition))
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$Trial_Index_)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$Trial_Index_)

```



```{r}
first4s.data.wider <- first4s.data %>%
  select(
    subject,
    session.num,
    condition_trial,
    Trial_Index_,
    IA_LABEL,
    IA_fam_object,
    condition,
    trial_type,
    agent_target_side_first,
    condition_first,
    video_file,
    DATA_FILE,
    IA_AREA,
    IA_DWELL_TIME,
    IA_FIXATION_COUNT
  ) %>% 
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  pivot_wider(names_from = IA_fam_object, values_from = )%>% #IA_DWELL_TIME for dwell times
  # mutate(object_congruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_lower_object, ifelse(target_location=="down",IA_upper_object, "") )), object_incongruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_upper_object, ifelse(target_location=="down",IA_lower_object, "") )))%>%
  # select(-IA_lower_object, -IA_upper_object, -IA_agent)%>%
  # pivot_longer(cols=c(object_congruent_dwell_time, object_incongruent_dwell_time), names_to = "object_congruent", values_to = "dwell_time")


#checks:
view(dfSummary(end.ip.data))

table(end.ip.data$trial_within_block, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$subject)
table(end.ip.data$session.num, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$TRIAL_TOTAL_VISITED_IA_COUNT)

```


```{r}

test.IP.end.data <- end.ip.data %>%
  group_by(subject, condition, target_location, IA_LABEL)%>%
  summarise(mean_dwell=mean(IA_DWELL_TIME))



ggplot(data = test.IP.end.data, aes(x=target_location, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  #geom_point(aes(x=first_target_loc, y=IA_DWELL_TIME, group=IA_LABEL),position_dodge(), alpha=0.5)+
  facet_grid(condition~IA_LABEL)+
  ylab("Dwell time")+xlab("Condition")+
  #ylim(0,3000)+
  theme_bw()#+
  #stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```

```{r}

test.IP.end.data.object.congruent <- end.ip.data.wider %>%
  group_by(subject, condition, object_congruent)%>%
  summarise(mean_dwell=mean(dwell_time))



ggplot(data = test.IP.end.data.object.congruent, aes(x=object_congruent, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  geom_point(alpha=0.3)+
  geom_line(aes(group=subject), alpha=0.3)+
  facet_wrap(~condition)+
  ylab("Dwell time")+xlab("Object congruency")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```


```{r}
test.IP.end.data.object.congruent.hand<-test.IP.end.data.object.congruent%>%
  filter(condition=="hand")

t.test(test.IP.end.data.object.congruent.hand$mean_dwell[test.IP.end.data.object.congruent.hand$object_congruent=="object_congruent_dwell_time"], test.IP.end.data.object.congruent.hand$mean_dwell[test.IP.end.data.object.congruent.hand$object_congruent=="object_incongruent_dwell_time"], paired=TRUE)
```




### End IP (longer)

```{r}
end.ip.long.orig.data <- read_delim("data/dog_woodward_paradigm_IA_report_video_IP_end_long.txt", na=".", delim="\t")

demo.data <- read_csv("data/Woodward_Kano_ counterbalancing.csv")%>%
  mutate(EDF1=fct_recode(as.factor(EDF1), "Georgia_1"="Georgia1"))%>%
  separate(EDF1, c("subject", "num"), sep="_")


end.ip.long.data <- end.ip.long.orig.data %>%
  select(Session_Name_, IP_LABEL, IA_LABEL,IA_ID, Trial_Index_, condition, target, target_location, session, trial_within_block, video_file,  video_latency, DATA_FILE,  IA_AREA, IA_AVERAGE_FIX_PUPIL_SIZE,  IA_DWELL_TIME, "IA_DWELL_TIME_%", IA_FIRST_FIXATION_TIME, IA_FIXATION_COUNT, IA_MAX_FIX_PUPIL_SIZE, INTEREST_AREA_FIXATION_SEQUENCE, TRIAL_DWELL_TIME, TRIAL_FIXATION_COUNT, TRIAL_IA_COUNT, TRIAL_TOTAL_VISITED_IA_COUNT )%>%
 #  filter(!(DATA_FILE=="Ace_2.edf" & Trial_Index_!=4))%>%
#  mutate(trial=ifelse(DATA_FILE=="Ace_2.edf" & Trial_Index_==4, 5, Trial_Index_))%>% #recoding of trial number to account for a dog that did not complete his first session, session was repeated with this dog
  mutate(Session_Name_=fct_recode(Session_Name_, Georgia_1="Georgia1",Georgia_2="Georgia2", Georgia_3="Georgia3",Georgia_4="Georgia4" ))%>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial=paste(condition, "_", trial_within_block))%>%
  full_join(demo.data)
  


end.ip.long.data.wider<-end.ip.long.data%>%
  select(subject, session.num,condition_trial, IP_LABEL, IA_LABEL, condition, target, target_location, session, trial_within_block, video_file,  video_latency, DATA_FILE,  IA_DWELL_TIME)%>%
  pivot_wider(names_from = IA_LABEL, values_from=IA_DWELL_TIME)%>%
  mutate(object_congruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_lower_object, ifelse(target_location=="down",IA_upper_object, "") )), object_incongruent_dwell_time=as.numeric(ifelse(target_location=="up",IA_upper_object, ifelse(target_location=="down",IA_lower_object, "") )))%>%
  select(-IA_lower_object, -IA_upper_object, -IA_agent)%>%
  pivot_longer(cols=c(object_congruent_dwell_time, object_incongruent_dwell_time), names_to = "object_congruent", values_to = "dwell_time")


#checks:
view(dfSummary(end.ip.data))

table(end.ip.data$trial_within_block, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$subject)
table(end.ip.data$session.num, end.ip.data$subject)
table(end.ip.data$condition, end.ip.data$TRIAL_TOTAL_VISITED_IA_COUNT)

```


```{r}

test.IP.end.long.data <- end.ip.long.data %>%
  group_by(subject, condition, target_location, IA_LABEL)%>%
  summarise(mean_dwell=mean(IA_DWELL_TIME))



ggplot(data = test.IP.end.long.data, aes(x=target_location, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  #geom_point(aes(x=first_target_loc, y=IA_DWELL_TIME, group=IA_LABEL),position_dodge(), alpha=0.5)+
  facet_grid(condition~IA_LABEL)+
  ylab("Dwell time")+xlab("Condition")+
  #ylim(0,3000)+
  theme_bw()#+
  #stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```

```{r}

test.IP.end.long.data.object.congruent <- end.ip.long.data.wider %>%
  group_by(subject, condition, object_congruent)%>%
  summarise(mean_dwell=mean(dwell_time))



ggplot(data = test.IP.end.long.data.object.congruent, aes(x=object_congruent, y=mean_dwell))+
  geom_boxplot(alpha=0.5)+#, outlier.colour = "white")+
  geom_point(alpha=0.3)+
  geom_line(aes(group=subject), alpha=0.3)+
  facet_wrap(~condition)+
  ylab("Dwell time")+xlab("Object congruency")+
  #ylim(0,3000)+
  theme_bw()+
  stat_compare_means(paired = TRUE, method = "t.test", label.y = 1500 )

```

### Trial report

```{r}

trialreport.end.ip.orig.data <- read_delim("data/dog_woodward_paradigm_trial_report_video_IP_end.txt", na=".", delim="\t")



```







-repear IA analysis with larger areas of interest
- setup a model for the dwell time (proportion of correct AOI)
- first look
- pupil size







############################################################

```{r}
#first test trial
test.IP.end.IA.endpos.trial1 <- end.ip.data %>%
  filter(condition!="hab", IA_LABEL=="IA_end_position", trial==4)


t.test(test.IP.end.IA.endpos.trial1$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="exp"], test.IP.end.IA.endpos.trial1$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="con"], paired=TRUE)

#second test trial
test.IP.end.IA.endpos.trial2 <- end.ip.data %>%
  filter(condition!="hab", IA_LABEL=="IA_end_position", trial==5)


t.test(test.IP.end.IA.endpos.trial2$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="exp"], test.IP.end.IA.endpos.trial2$IA_DWELL_TIME[test.IP.end.IA.endpos.trial1$condition=="con"], paired=TRUE)


```


### Onscreen Dwell time (while videos is playing)

```{r}
video_playing_data<-read.delim("data/ET_VoE_occlusion2screens_IA_report_whole_screen_video_playing_IP.txt",na=".", sep = "\t")%>%
  select(Session_Name_, IP_LABEL, IA_LABEL,IA_ID, Trial_Index_, condition, phase, session,  video_file, video_id, video_latency, DATA_FILE,  IA_AREA, IA_AVERAGE_FIX_PUPIL_SIZE,  IA_DWELL_TIME, "IA_DWELL_TIME", IA_FIXATION_COUNT, IA_MAX_FIX_PUPIL_SIZE, INTEREST_AREA_FIXATION_SEQUENCE, TRIAL_DWELL_TIME, TRIAL_FIXATION_COUNT, TRIAL_IA_COUNT, TRIAL_TOTAL_VISITED_IA_COUNT )%>%
   filter(!(DATA_FILE=="Ace_2.edf" & Trial_Index_!=4))%>%
  mutate(trial=ifelse(DATA_FILE=="Ace_2.edf" & Trial_Index_==4, 5, Trial_Index_))%>% #recoding of trial number to account for a dog that did not complete his first session, session was repeated with this dog
  mutate(Session_Name_=fct_recode(Session_Name_, Georgia_1="Georgia1",Georgia_2="Georgia2", Cheynna_1="Cheynna1", Cheynna_2="Cheynna2" ))%>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial=paste(condition, "_", trial))%>%
  full_join(demo.data)
 # rename(trial="Trial_Index_")
  

#checks:
view(dfSummary(video_playing_data))


on_screen_data<-video_playing_data%>%
  filter(IA_LABEL=="screen_aoi")%>%
  mutate(prop_looking=IA_DWELL_TIME/5000)


on_screen_data%>%group_by(condition)%>%
  summarise(mean=mean(prop_looking, na.rm = TRUE), median=median(prop_looking, na.rm = TRUE), se=sd(prop_looking, na.rm = TRUE)/sqrt(length(prop_looking)), min=min(prop_looking, na.rm = TRUE), max=max(prop_looking, na.rm = TRUE))

```

onscreen dwell-time: comparison between first and last familiarisation trial.
```{r}
library(broom)
on_screen_data%>%filter(phase=="hab1" | phase=="hab3")%>%
  group_by(subject, phase)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ phase, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```


onscreen dwell-time: comparison between test trials
* first test trial
```{r}
on_screen_data%>%filter(condition!="hab", trial==4)%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ condition, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```

* second test trial
```{r}
on_screen_data%>%filter(condition!="hab", trial==5)%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking, na.rm=TRUE))%>%
  ungroup()%>%
  summarise(ttest = list(t.test(mean ~ condition, paired=TRUE)))%>%
  mutate(ttest = map(ttest, tidy)) %>%
  unnest()
```

Plot onscreen dwell time in test trials
```{r}
xx<-on_screen_data%>%filter(condition!="hab")%>%
  group_by(subject, condition)%>%
  summarise(mean=mean(prop_looking))%>%
  ungroup()

ggplot(data=xx, aes(x=condition, y=mean))+
  geom_boxplot()
```