---
title: Woodward 2022: Dwell time analysis
author: Christoph Voelter
date: 
output: html_document
---
notes: 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
#remove(list=c("first4s.data_short"))
library(tidyverse)
library(summarytools)
library(lme4)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
#load("workspaces/AL_test_first4s.RData")
```

## Original IAs
### End IP
```{r}
demo.data <- read_csv("data/Woodward_2022_counterbalancing.csv") %>%
  mutate(
    EDF1 = fct_recode(
      as.factor(EDF1),
      "Georgia_1" = "Georgia1",
    )
  ) %>%
  separate(EDF1, c("subject", "num"), sep = "_")
```

```{r}
after_agent_dir_cue.data <-
  read_delim(
    "data/Woodward2022_IA_Report_for_combined-agent-object_analysis_IP_after_agent_directional_cue.txt",
    na = ".",
    delim = "\t"
  ) %>%
  dplyr::select(
    Session_Name_,
    IP_LABEL,
    IA_LABEL,
    IA_ID,
    Trial_Index_,
    condition,
    dog_session,
    sub_id,
    trial_number,
    trial_type,
    session_number,
    video_file,
    DATA_FILE,
    IA_AREA,
    IA_AVERAGE_FIX_PUPIL_SIZE,
    IA_DWELL_TIME,
    "IA_DWELL_TIME_%",
    IA_FIXATION_COUNT,
    "IA_FIXATION_%",
    IA_FIRST_FIXATION_INDEX, #Ordinal sequence of the first fixation that       #was within the current interest area.
    IA_FIRST_FIXATION_VISITED_IA_COUNT, #number of different interest areas
    #visited so far before the first fixation is made to the
    #current interest area.
    IA_MAX_FIX_PUPIL_SIZE,
    IA_FIRST_FIXATION_TIME,
    INTEREST_AREA_FIXATION_SEQUENCE,
    TRIAL_DWELL_TIME,
    TRIAL_FIXATION_COUNT,
    TRIAL_IA_COUNT,
    TRIAL_TOTAL_VISITED_IA_COUNT
  ) %>%
  filter(DATA_FILE != "Melody_1.edf", DATA_FILE != "Mia_5.edf") %>%
  filter(DATA_FILE != "Timo_5.edf", DATA_FILE != "Joker_5.edf") %>% # backup recordings in case the first sessions are below 70% onscreen looking
  #filter(!(DATA_FILE == "Amy_4.edf" & Trial_Index_ != 1)) %>%
  #mutate(session.num = as.numeric(session.num)) %>%
  #mutate(session.num.corrected = ifelse(DATA_FILE == "Amy_4.edf" | DATA_FILE == "Amy_5.edf" | DATA_FILE == "George_5.edf" | DATA_FILE == "Melody_2.edf" | DATA_FILE == #"Melody_3.edf" | DATA_FILE == "Melody_4.edf", session.num-1,
#                                        ifelse(DATA_FILE == "Melody_7.edf" | DATA_FILE == "Mia_6.edf", session.num-3, session.num))) %>% #recoding of session number to account for dogs with aborted sessions
  mutate(
    Session_Name_ = fct_recode(
      Session_Name_,
      Georgia_1 = "Georgia1",
      Georgia_2 = "Georgia2",
      Georgia_3 = "Georgia3",
      Georgia_4 = "Georgia4"
    )
  ) %>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial = paste(condition, "_", Trial_Index_)) %>%
  filter(condition != "fam")%>% #remove familiarisation 
  full_join(demo.data) %>%
  mutate(IA_LABEL2 = IA_LABEL, trial_type2 = trial_type) %>%
  separate(
    IA_LABEL2,
    into = c("ia_object", "ia_side", "ia_label_del", "ia_label_del2"),
    sep = "_"
  ) %>%
  separate(
    trial_type2,
    into = c("trialtype_agent", "trialtype_object", "trialtype_side"),
    sep = "_"
  ) %>%
  #filter(ia_object == "elephant" | ia_object == "ball") %>%
  mutate(
    IA_fam_object = ifelse(
      condition == "new_goal" &
        ia_object == trialtype_object,
      "new_goal_old_side",
      ifelse(
        condition == "new_goal" &
          ia_object != trialtype_object,
        "old_goal_new_side",
        ifelse(
          condition == "new_side" &
            ia_object != trialtype_object,
          "new_goal_old_side",
          ifelse(
            condition == "new_side" &
              ia_object == trialtype_object,
            "old_goal_new_side",
            ""
          )
        )
      )
    )
  ) %>%
  filter(is.na(ia_label_del2)) %>% #remove the smaller IAs from familiarisation
  mutate(ia_object = ifelse(trialtype_agent =="inanimate" & ia_object == "human", "inanimate", ia_object))%>%
  separate(
    agent_target_side_first,
    into = c("agent_first", "object_first", "side_first"), #shown to the dog in the first session
    sep = "_"
  ) 
  
after_agent_dir_cue.data %>% select(subject, condition_trial, trial_type, IA_LABEL, IA_fam_object) 
#IA_fam_object=what the AOI looked at contains
after_agent_dir_cue.data %>% select(subject, condition, trial_type, IA_LABEL, IA_fam_object,IA_FIRST_FIXATION_INDEX,IA_FIRST_FIXATION_VISITED_IA_COUNT) 


levels(as.factor(after_agent_dir_cue.data$DATA_FILE))
levels(as.factor(after_agent_dir_cue.data$condition))
table(as.factor(after_agent_dir_cue.data$DATA_FILE),
      after_agent_dir_cue.data$Trial_Index_)
table(as.factor(after_agent_dir_cue.data$DATA_FILE),
      after_agent_dir_cue.data$Trial_Index_)
table(as.factor(after_agent_dir_cue.data$DATA_FILE),
      after_agent_dir_cue.data$session.num)
table(as.factor(after_agent_dir_cue.data$DATA_FILE),
      after_agent_dir_cue.data$session_number)
```

```{r}

after_agent_dir_cue_wider.data <- after_agent_dir_cue.data %>%
  filter(ia_object == "human" | ia_object == "inanimate" | ia_object == trialtype_object) %>% #only keep agent IA and active
  select(subject, session_number, ia_object, TRIAL_DWELL_TIME, condition, agent_first, ,trial_type, trialtype_agent, trialtype_object, trialtype_side,  IA_DWELL_TIME)%>%
  group_by(subject, session_number, condition, agent_first, TRIAL_DWELL_TIME, trial_type, trialtype_agent, trialtype_object, trialtype_side) %>%
  summarise(sum_dwell_time = sum(IA_DWELL_TIME))%>%
  ungroup()%>%
  mutate(prop_dwell_time = sum_dwell_time / TRIAL_DWELL_TIME)
```
## Linear mixed model

```{r}

after_agent_dir_cue_wider.data$z.session<-as.vector(scale(after_agent_dir_cue_wider.data$session_number, center = TRUE, scale=TRUE))

after_agent_dir_cue_wider.data$condition <-
  as.factor(after_agent_dir_cue_wider.data$condition)
levels(after_agent_dir_cue_wider.data$condition)
after_agent_dir_cue_wider.data$condition.c = as.vector(scale(
  as.numeric(
    after_agent_dir_cue_wider.data$condition == levels(as.factor(
      after_agent_dir_cue_wider.data$condition
    ))[2]
  ),
  center = TRUE,
  scale = FALSE
))


after_agent_dir_cue_wider.data$trialtype_agent <-
  as.factor(after_agent_dir_cue_wider.data$trialtype_agent)
levels(after_agent_dir_cue_wider.data$trialtype_agent)

after_agent_dir_cue_wider.data$trialtype_agent.c = as.vector(scale(
  as.numeric(
    after_agent_dir_cue_wider.data$trialtype_agent == levels(as.factor(after_agent_dir_cue_wider.data$trialtype_agent))[2]
  ),
  center = TRUE,
  scale = FALSE
))

after_agent_dir_cue_wider.data$agent_first.c = as.vector(scale(
  as.numeric(after_agent_dir_cue_wider.data$agent_first == levels(
    as.factor(after_agent_dir_cue_wider.data$agent_first)
  )[2]),
  center = TRUE,
  scale = FALSE
))
levels(as.factor(after_agent_dir_cue_wider.data$z.session))
hist(after_agent_dir_cue_wider.data$sum_dwell_time)
```




fit model
```{r}

mm1.dwell.time<-lme4::lmer(sum_dwell_time~ condition * trialtype_agent +
                            z.session +  agent_first +
            (1+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=FALSE)
#does not converge
mm1.dwell.time_red<-lme4::lmer(sum_dwell_time~ condition * trialtype_agent +
                            z.session +  agent_first +
            (1|subject)+ (0+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=FALSE)


drop1(mm1.dwell.time_red, test="Chisq")
# summary(mm1.yarbus.signaling)
```


```{r}
diagnostics.plot(mm1.dwell.time_red, size.fac=2)

ranef.diagn.plot(mm1.dwell.time_red)
```

```{r}
# drop1_mm1.yarbus.signaling <- drop1(mm1.yarbus.signaling, test="Chisq")%>%
#   filter(!is.na(npar))%>%
#   add_row(npar = rep(NA,1),  .before = 1)%>%
#   add_row(npar  = rep(NA,3),  .before = 3)
# drop1_mm1.yarbus.signaling

```
Individual fixed effects
```{r}
library(lmerTest)

mm1.dwell.time_red.reml<-lmerTest::lmer(sum_dwell_time~ condition * trialtype_agent +
                            z.session +  agent_first +
            (1|subject)+ (0+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=TRUE)

summary(mm1.dwell.time_red.reml)
```
 ### reduced model without interaction term
```{r}

mm1.dwell.time_red2.reml<-lmerTest::lmer(sum_dwell_time~ condition + trialtype_agent +
                            z.session +  agent_first +
            (1|subject)+ (0+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=TRUE)

summary(mm1.dwell.time_red2.reml)

```
# until here 
 
Colinearity checks
```{r}
library(car)
xx=lm(baitedX_corrected~condition + z.trial+z.session + sex + z.age, data=yarbus.data.signaling)
vif(xx)
```
relative model complexity
```{r}
length(residuals(mm1.dwell.time_red2.reml))/
(length(fixef(mm1.dwell.time_red2.reml))+
nrow(as.data.frame(summary(mm1.dwell.time_red2.reml)$varcor)))
```
model stability
```{r}

mm1.yarbus.signaling.stab=glmm.model.stab(model.res=mm1.yarbus.signaling, contr=NULL, para=F, data=NULL)

mm1.yarbus.signaling.stab$summary

m.stab.plot(round(mm1.yarbus.signaling.stab$summary[, -1], 3))
```
--> model stable with regard to the fixed effects

```{r}
boot.mm1.yarbus.signaling=boot.lmer(mm1.yarbus.signaling, discard.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)

mm1.yarbus.signaling_boot_ci<-boot.mm1.yarbus.signaling$ci.estimates
```
#### effect size
```{r}
library(MuMIn)
r.squaredGLMM(mm1.yarbus.signaling)
```


#### output table
```{r}


model_table_mm1.yarbus.signaling <- bind_cols(as.data.frame(summary(mm1.yarbus.signaling)$coefficients),
                         drop1_mm1.yarbus.signaling,
                         mm1.yarbus.signaling_boot_ci) %>% #mm2_boot_ci<-boot.full$ci.estimates
  dplyr::select(Estimate, SE = `Std. Error`, LowerCI = X2.5., UpperCI = X97.5., Chi2 = LRT, df = npar, p = `Pr(Chi)`) %>%
  mutate(across(.cols = c(p), ~ round(.x, 3))) %>% 
  mutate(across(.cols = c(Estimate:Chi2), ~ round(.x, 2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
mutate(p=replace(p, p==0, "<0.001"))

write.csv(model_table_mm1.yarbus.signaling , file = "saves/yarbus_signalling_mm1_output_table.csv")
```




#######beta model, old code########


## fit model
```{r}
mm1_signalling.face=glmmTMB(prop_fixation_time.scaled ~ 
                             condition+z.trial+z.session+z.age+sex +
                             (1|subject)+
                             (0+condition.c1|subject)+(0+condition.c2|subject)+ (0+condition.c3|subject)+(0+condition.c4|subject)+
                             (0+z.session|subject),# + (0+z.trial|subject), pruned to convergence issues
                           data=signalling.face.data, family=beta_family, control=contr)

summary(mm1_signalling.face)
```


```{r}
mm1_signalling.face_null=glmmTMB(prop_fixation_time.scaled ~ 
                             z.trial+z.session+z.age+sex +
            (1|subject)+
            (0+condition.c1|subject)+(0+condition.c2|subject)+ (0+condition.c3|subject)+(0+condition.c4|subject)+
            (0+z.session|subject),# + (0+z.trial|subject),
             data=signalling.face.data, family=beta_family, control=contr)

anova(mm1_signalling.face, mm1_signalling.face_null, test="Chisq")

```

LRT
```{r}
mm1_signalling.face_drop1 <- drop1(mm1_signalling.face, test="Chisq")
mm1_signalling.face_drop1
mm1_signalling.face_drop1<- mm1_signalling.face_drop1%>% 
  filter(!is.na(Df)) %>% 
  add_row(Df = rep(NA,1),  .before = 1) %>% 
  add_row(Df = rep(NA,3),  .before = 3)
```

+ check for overdispersion
```{r}
overdisp.test(mm1_signalling.face)
```


+ Collinearity checks
```{r}
library(car)
xx=lm(prop_fixation_time.scaled ~                              condition+z.trial+z.session+z.age+sex, data=signalling.face.data)
vif(xx)
```

```{r eval=FALSE}
mm1_signalling.face.ci=boot.glmmtmb(mm1_signalling.face, 
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95, data=signalling.face.data)
```

#### output table

```{r}
mm1_signalling.face_table <- bind_cols(as.data.frame(summary(mm1_signalling.face)$coefficients$cond),
                                      mm1_signalling.face_drop1)%>%#,
                                    # mm1_area_tail.ci$ci.estimates$fe[1:7,]) %>%
  dplyr::select(Estimate, SE = `Std. Error`, Chi2 = LRT, df = Df, p = `Pr(>Chi)`, z_wald=`z value`, p_wald=`Pr(>|z|)`) %>% #LowerCI = X2.5., UpperCI = X97.5., 
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall=3))) %>% 
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall=2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p=replace(p, p==0, "<0.001"))

write.csv(mm1_signalling.face_table, file = "saves/mm1_signalling.face_table.csv")
```

