---
title: Woodward 2022: Dwell time analysis
author: Christoph Voelter
date: 
output: html_document
---
notes: 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
#remove(list=c("first4s.data_short"))
library(tidyverse)
library(summarytools)
library(lme4)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
load("./workspaces/Looking_times.RData")
```

## Import the data
```{r}
demo.data <- read_csv("data/Woodward_2022_counterbalancing.csv") %>%
  mutate(
    EDF1 = fct_recode(
      as.factor(EDF1),
      "Georgia_1" = "Georgia1",
    )
  ) %>%
  separate(EDF1, c("subject", "num"), sep = "_")
```

```{r}
after_agent_dir_cue.data <-
  read_delim(
    "data/Woodward2022_IA_Report_for_combined-agent-object_analysis_IP_after_agent_directional_cue.txt",
    na = ".",
    delim = "\t"
  ) %>%
  dplyr::select(
    Session_Name_,
    IP_LABEL,
    IA_LABEL,
    IA_ID,
    Trial_Index_,
    condition,
    dog_session,
    sub_id,
    trial_number,
    trial_type,
    session_number,
    video_file,
    DATA_FILE,
    IA_AREA,
    IA_AVERAGE_FIX_PUPIL_SIZE,
    IA_DWELL_TIME,
    "IA_DWELL_TIME_%",
    IA_FIXATION_COUNT,
    "IA_FIXATION_%",
    IA_FIRST_FIXATION_INDEX, #Ordinal sequence of the first fixation that       #was within the current interest area.
    IA_FIRST_FIXATION_VISITED_IA_COUNT, #number of different interest areas
    #visited so far before the first fixation is made to the
    #current interest area.
    IA_MAX_FIX_PUPIL_SIZE,
    IA_FIRST_FIXATION_TIME,
    INTEREST_AREA_FIXATION_SEQUENCE,
    TRIAL_DWELL_TIME,
    TRIAL_FIXATION_COUNT,
    TRIAL_IA_COUNT,
    TRIAL_TOTAL_VISITED_IA_COUNT
  ) %>%
  filter(DATA_FILE != "Melody_1.edf", DATA_FILE != "Mia_5.edf") %>%
  filter(DATA_FILE != "Timo_5.edf", DATA_FILE != "Joker_5.edf") %>% # backup recordings in case the first sessions are below 70% onscreen looking
  #filter(!(DATA_FILE == "Amy_4.edf" & Trial_Index_ != 1)) %>%
  #mutate(session.num = as.numeric(session.num)) %>%
  #mutate(session.num.corrected = ifelse(DATA_FILE == "Amy_4.edf" | DATA_FILE == "Amy_5.edf" | DATA_FILE == "George_5.edf" | DATA_FILE == "Melody_2.edf" | DATA_FILE == #"Melody_3.edf" | DATA_FILE == "Melody_4.edf", session.num-1,
#                                        ifelse(DATA_FILE == "Melody_7.edf" | DATA_FILE == "Mia_6.edf", session.num-3, session.num))) %>% #recoding of session number to account for dogs with aborted sessions
  mutate(
    Session_Name_ = fct_recode(
      Session_Name_,
      Georgia_1 = "Georgia1",
      Georgia_2 = "Georgia2",
      Georgia_3 = "Georgia3",
      Georgia_4 = "Georgia4"
    )
  ) %>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial = paste(condition, "_", Trial_Index_)) %>%
  filter(condition != "fam")%>% #remove familiarisation 
  full_join(demo.data) %>%
  mutate(IA_LABEL2 = IA_LABEL, trial_type2 = trial_type) %>%
  separate(
    IA_LABEL2,
    into = c("ia_object", "ia_side", "ia_label_del", "ia_label_del2"),
    sep = "_"
  ) %>%
  separate(
    trial_type2,
    into = c("trialtype_agent", "trialtype_object", "trialtype_side"),
    sep = "_"
  ) %>%
  #filter(ia_object == "elephant" | ia_object == "ball") %>%
  mutate(
    IA_fam_object = ifelse(
      condition == "new_goal" &
        ia_object == trialtype_object,
      "new_goal_old_side",
      ifelse(
        condition == "new_goal" &
          ia_object != trialtype_object,
        "old_goal_new_side",
        ifelse(
          condition == "new_side" &
            ia_object != trialtype_object,
          "new_goal_old_side",
          ifelse(
            condition == "new_side" &
              ia_object == trialtype_object,
            "old_goal_new_side",
            ""
          )
        )
      )
    )
  ) %>%
  filter(is.na(ia_label_del2)) %>% #remove the smaller IAs from familiarisation
  mutate(ia_object = ifelse(trialtype_agent =="inanimate" & ia_object == "human", "inanimate", ia_object))%>%
  separate(
    agent_target_side_first,
    into = c("agent_first", "object_first", "side_first"), #shown to the dog in the first session
    sep = "_"
  ) 
  
after_agent_dir_cue.data %>% select(subject, condition_trial, trial_type, IA_LABEL, IA_fam_object) 
#IA_fam_object=what the AOI looked at contains
after_agent_dir_cue.data %>% select(subject, condition, trial_type, IA_LABEL, IA_fam_object,IA_FIRST_FIXATION_INDEX,IA_FIRST_FIXATION_VISITED_IA_COUNT) 


levels(as.factor(after_agent_dir_cue.data$DATA_FILE))
levels(as.factor(after_agent_dir_cue.data$condition))
table(as.factor(after_agent_dir_cue.data$DATA_FILE),
      after_agent_dir_cue.data$Trial_Index_)
table(as.factor(after_agent_dir_cue.data$DATA_FILE),
      after_agent_dir_cue.data$Trial_Index_)
table(as.factor(after_agent_dir_cue.data$DATA_FILE),
      after_agent_dir_cue.data$session.num)
table(as.factor(after_agent_dir_cue.data$DATA_FILE),
      after_agent_dir_cue.data$session_number)
```

```{r}

after_agent_dir_cue_wider.data <- after_agent_dir_cue.data %>%
  filter(ia_object == "human" | ia_object == "inanimate" | ia_object == trialtype_object) %>% #only keep agent IA and active
  select(subject, session_number, ia_object, TRIAL_DWELL_TIME, condition, agent_first, ,trial_type, trialtype_agent, trialtype_object, trialtype_side,  IA_DWELL_TIME)%>%
  group_by(subject, session_number, condition, agent_first, TRIAL_DWELL_TIME, trial_type, trialtype_agent, trialtype_object, trialtype_side) %>%
  summarise(sum_dwell_time = sum(IA_DWELL_TIME))%>%
  ungroup()%>%
  mutate(prop_dwell_time = sum_dwell_time / TRIAL_DWELL_TIME)
```

## Linear mixed model

```{r}

after_agent_dir_cue_wider.data$z.session<-as.vector(scale(after_agent_dir_cue_wider.data$session_number, center = TRUE, scale=TRUE))

after_agent_dir_cue_wider.data$condition <-
  as.factor(after_agent_dir_cue_wider.data$condition)
levels(after_agent_dir_cue_wider.data$condition)
after_agent_dir_cue_wider.data$condition.c = as.vector(scale(
  as.numeric(
    after_agent_dir_cue_wider.data$condition == levels(as.factor(
      after_agent_dir_cue_wider.data$condition
    ))[2]
  ),
  center = TRUE,
  scale = FALSE
))


after_agent_dir_cue_wider.data$trialtype_agent <-
  as.factor(after_agent_dir_cue_wider.data$trialtype_agent)
levels(after_agent_dir_cue_wider.data$trialtype_agent)

after_agent_dir_cue_wider.data$trialtype_agent.c = as.vector(scale(
  as.numeric(
    after_agent_dir_cue_wider.data$trialtype_agent == levels(as.factor(after_agent_dir_cue_wider.data$trialtype_agent))[2]
  ),
  center = TRUE,
  scale = FALSE
))

after_agent_dir_cue_wider.data$agent_first.c = as.vector(scale(
  as.numeric(after_agent_dir_cue_wider.data$agent_first == levels(
    as.factor(after_agent_dir_cue_wider.data$agent_first)
  )[2]),
  center = TRUE,
  scale = FALSE
))
levels(as.factor(after_agent_dir_cue_wider.data$z.session))
hist(after_agent_dir_cue_wider.data$sum_dwell_time)
```




fit model
```{r}

mm1.dwell.time<-lme4::lmer(sum_dwell_time~ condition * trialtype_agent +
                            z.session +  agent_first +
            (1+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=FALSE)
#does not converge
mm1.dwell.time_red<-lme4::lmer(sum_dwell_time~ condition * trialtype_agent +
                            z.session +  agent_first +
            (1|subject)+ (0+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=FALSE)
```


```{r}
diagnostics.plot(mm1.dwell.time_red, size.fac=2)

ranef.diagn.plot(mm1.dwell.time_red)
```

Individual fixed effects
```{r}
summary(mm1.dwell.time_red)
library(lmerTest)

mm1.dwell.time_red.reml<-lmerTest::lmer(sum_dwell_time ~ condition * trialtype_agent +
                            z.session +  agent_first +
            (1|subject)+ (0+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=TRUE)
summary(mm1.dwell.time_red.reml)
```
 ### reduced model without interaction term
```{r}
mm1.dwell.time_red2<-lme4::lmer(sum_dwell_time ~ condition + trialtype_agent +
                            z.session +  agent_first +
            (1|subject)+ (0+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=FALSE)
#null model lacking the fixed effects of condition and agent
mm1.swell.time_red2_null<-lme4::lmer(sum_dwell_time~ 
                            z.session +  agent_first +
            (1|subject)+ (0+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=FALSE)

#full-null model comparison
as.data.frame(anova(mm1.swell.time_red2_null,mm1.dwell.time_red2, test="Chisq"))#signif.

#Individual fixed effects
mm1.dwell.time_red2.reml<-lmerTest::lmer(sum_dwell_time~ condition + trialtype_agent +
                            z.session +  agent_first +
            (1|subject)+ (0+condition.c+trialtype_agent.c + z.session|subject), data=after_agent_dir_cue_wider.data, REML=TRUE)

summary(mm1.dwell.time_red2.reml) #irrespective of condition, longer lt to the human agent

```
 
Collinearity checks
```{r}
library(car)
xx=lm(sum_dwell_time ~ condition + trialtype_agent +
           z.session +  agent_first, 
           data=after_agent_dir_cue_wider.data)
vif(xx)
```
relative model complexity
```{r}
length(residuals(mm1.dwell.time_red2))/
(length(fixef(mm1.dwell.time_red2))+
nrow(as.data.frame(summary(mm1.dwell.time_red2)$varcor)))
```
Descriptive statistics
```{r}
after_agent_dir_cue_wider.data.descr <- after_agent_dir_cue_wider.data %>% 
  group_by(trialtype_agent) %>% 
  summarise(mean_dwell_time = mean(sum_dwell_time),
            sd_dwell_time = sd(sum_dwell_time))

after_agent_dir_cue_wider.data.descr
```

model stability
```{r}

mm1.dwell.time_red2.stab=glmm.model.stab(model.res=mm1.dwell.time_red2, contr=NULL, para=F, data=NULL)

mm1.dwell.time_red2.stab$summary

png("graphs/mm1_dwell_time_red2_stability_plot.png")
m.stab.plot(round(mm1.dwell.time_red2.stab$summary[, -1], 3))
dev.off()


```
--> model stable with regard to the fixed effects

confidence intervals
```{r}
boot.mm1.dwell.time_red2=boot.glmm.pred(mm1.dwell.time_red2, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)

mm1.dwell.time_red2_boot_ci_est<-boot.mm1.dwell.time_red2$ci.estimates
mm1.dwell.time_red2_boot_ci_est
```
#### effect size
```{r}
library(MuMIn)
r.squaredGLMM(mm1.dwell.time_red2)
```


#### output table
```{r}
model_table_mm1.dwell.time_red2 <- bind_cols(as.data.frame(summary(mm1.dwell.time_red2)$coefficients[,-3]),
                         summary(mm1.dwell.time_red2.reml)$coefficients[,c("df","t value", "Pr(>|t|)")],
                         mm1.dwell.time_red2_boot_ci_est[,2:3],
                         mm1.dwell.time_red2.stab$summary[1:5, -(1:2)]) %>% 
  dplyr::select(Estimate, SE = `Std. Error`, t = `t value`, df, LowerCI = X2.5., UpperCI = X97.5., P = `Pr(>|t|)`, min, max) %>%
  mutate(across(.cols = c(P), ~ round(.x, 3))) %>% 
  mutate(across(.cols = -"P", ~ round(.x, 2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
mutate(P=replace(P, P==0, "<0.001"))

write.csv(model_table_mm1.dwell.time_red2 , file = "saves/mm1.dwell.time_red2_agent_obj_test_output_table.csv")
```
## Plot absolute dwell times in the two conditions
```{r}
library(ggbeeswarm)

after_agent_dir_cue_wider.data.agg<- after_agent_dir_cue_wider.data %>%
  group_by(condition, subject)%>%
  summarise(mean_dwell_time = mean(sum_dwell_time))#averaged across trials and agents

lt_plot_cond<-ggplot(after_agent_dir_cue_wider.data.agg%>%
               mutate(condition= fct_recode(condition, "Old side-new identity"="new_goal", "Old identity-new side" = "new_side")),
                aes(x=condition,y=mean_dwell_time))+
                geom_boxplot(position = position_dodge(0.8), alpha=0.8)+
                geom_quasirandom(method = "pseudorandom", width=0.1, dodge.width=.8, alpha=.4, size=2)+
                geom_line(aes(group=subject,condition), linewidth=.5, lty =2, color="grey", alpha=.9)+
                ylab("Mean dwell time to agent/target obejct (ms)")+
                xlab("Condition")+
                ylim(c(1500, 18000))+
                theme_bw()+
                theme(axis.text = element_text(size = 11))+
                theme(axis.title = element_text(size = 14))
                    
            
                
lt_plot_cond
```

```{r}
ggsave(filename="graphs/looking_times.png",
  plot = lt_plot,
  width=8.5,
  height=6,
  #scale=0.55,
  dpi=1200)
```
## Plot absolute dwell times to the different agents
```{r}
library(ggsignif)

after_agent_dir_cue_wider.data.agg<- after_agent_dir_cue_wider.data %>%
  group_by(trialtype_agent, subject)%>%
  summarise(mean_dwell_time = mean(sum_dwell_time))#points averaged across trials and conditions

lt_plot_agent<-ggplot(after_agent_dir_cue_wider.data.agg%>%
              mutate(trialtype_agent= fct_recode(trialtype_agent, "Human"="human","Inanimate"="inanimate")),
                aes(x=trialtype_agent,y=mean_dwell_time))+
                geom_boxplot(position = position_dodge(0.8), alpha=0.8)+
  geom_quasirandom(method = "pseudorandom", width=0.1, dodge.width=.8, alpha=.4, size=2)+
                geom_line(aes(group=subject), linewidth=.5, lty =2, color="grey", alpha=.9)+
                ylab("Mean dwell time to agent/target obejct (ms)")+
                xlab("Agent")+
                ylim(c(1500, 18000))+
                geom_signif(comparisons=list(c("Human","Inanimate")), annotations="**",
                y_position = 16000, tip_length = 0, textsize=8)+
                theme_bw()+
                theme(axis.text = element_text(size = 11))+
                theme(axis.title = element_text(size = 14))
            
                
lt_plot_agent
```

```{r}
ggsave(filename="graphs/looking_times_agents.png",
  plot = lt_plot_agent,
  width=8.5,
  height=5,
  #scale=0.55,
  dpi=1200)
```
##Plot dwell times together
```{r}
library(cowplot)
bothplots<-plot_grid(lt_plot_cond, lt_plot_agent, nrow=1, labels=c("A", "B"),
          hjust = 0, vjust= 1.7, label_x = c(0.27,0.27))
bothplots
```
```{r}
ggsave(filename="graphs/dwell_times_Test_Exp2_both_plots.png",
  plot = bothplots,
  width=8.5,
  height=5,
  #scale=0.55,
  dpi=1200)
```

