---
title: Woodward 2022: Dwell time analysis
author: Christoph Voelter
date: 
output: html_document
---
notes: 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#rm(list=ls())
#remove(list=c("first4s.data_short"))
install.packages('TMB', type = 'source')

library(tidyverse)
library(summarytools)
library(lme4)
library(wesanderson)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
source("./functions/glmmTMB_stability.r")
source("./functions/boot_glmmTMB.r")
source("./functions/extract_ranef_gmmTMB.r")
load("workspaces/AL_test_first4s.RData")
```

```{r}
load("workspaces/AL_test_first4s.RData")
```


```{r}
demo.data <- read_csv("data/Woodward_2022_counterbalancing.csv") %>%
  mutate(
    EDF1 = fct_recode(
      as.factor(EDF1),
      "Georgia_1" = "Georgia1",
    )
  ) %>%
  separate(EDF1, c("subject", "num"), sep = "_")
```

```{r}
first4s.data <-
  read_delim(
    "data/Woodward2022_IA_Report_IP_test_first_4_sec.txt",
    na = ".",
    delim = "\t"
  ) %>%
  dplyr::select(
    Session_Name_,
    IP_LABEL,
    IA_LABEL,
    IA_ID,
    Trial_Index_,
    condition,
    dog_session,
    sub_id,
    trial_number,
    trial_type,
    session_number,
    video_file,
    DATA_FILE,
    IA_AREA,
    IA_AVERAGE_FIX_PUPIL_SIZE,
    IA_DWELL_TIME,
    "IA_DWELL_TIME_%",
    IA_FIXATION_COUNT,
    "IA_FIXATION_%",
    IA_FIRST_FIXATION_INDEX, #Ordinal sequence of the first fixation that       #was within the current interest area.
    IA_FIRST_FIXATION_VISITED_IA_COUNT, #number of different interest areas
    #visited so far before the first fixation is made to the
    #current interest area.
    IA_MAX_FIX_PUPIL_SIZE,
    IA_FIRST_FIXATION_TIME,
    INTEREST_AREA_FIXATION_SEQUENCE,
    TRIAL_DWELL_TIME,
    TRIAL_FIXATION_COUNT,
    TRIAL_IA_COUNT,
    TRIAL_TOTAL_VISITED_IA_COUNT,
    IP_START_TIME,
    TRIAL_START_TIME
  ) %>%
  filter(DATA_FILE != "Melody_1.edf", DATA_FILE != "Mia_5.edf") %>%
  filter(DATA_FILE != "Timo_5.edf", DATA_FILE != "Joker_5.edf") %>% # backup recordings in case the first sessions are below 70% onscreen looking
  mutate(
    Session_Name_ = fct_recode(
      Session_Name_,
      Georgia_1 = "Georgia1",
      Georgia_2 = "Georgia2",
      Georgia_3 = "Georgia3",
      Georgia_4 = "Georgia4"
    )
  ) %>%
  separate(Session_Name_, c("subject", "session.num"), sep = "_") %>%
  mutate(condition_trial = paste(condition, "_", Trial_Index_)) %>%
  full_join(demo.data) %>%
  mutate(IA_LABEL2 = IA_LABEL, trial_type2 = trial_type) %>%
  separate(
    IA_LABEL2,
    into = c("ia_object", "ia_side", "ia_label_del", "ia_label_del2"),
    sep = "_"
  ) %>%
  separate(
    trial_type2,
    into = c("trialtype_agent", "trialtype_object", "trialtype_side"),
    sep = "_"
  ) %>%
  filter(ia_object == "elephant" | ia_object == "ball") %>%
  mutate(
    IA_fam_object = ifelse(
      condition == "new_goal" &
        ia_object == trialtype_object,
      "new_goal_old_side",
      ifelse(
        condition == "new_goal" &
          ia_object != trialtype_object,
        "old_goal_new_side",
        ifelse(
          condition == "new_side" &
            ia_object != trialtype_object,
          "new_goal_old_side",
          ifelse(
            condition == "new_side" &
              ia_object == trialtype_object,
            "old_goal_new_side",
            ""
          )
        )
      )
    )
  ) %>%
  mutate(attention_getter_duration = IP_START_TIME - TRIAL_START_TIME)

first4s.data %>% select(subject, condition_trial, trial_type, IA_LABEL, IA_fam_object) 
#IA_fam_object=what the AOI looked at contains
first4s.data %>% select(subject, condition, trial_type, IA_LABEL, IA_fam_object,IA_FIRST_FIXATION_INDEX,IA_FIRST_FIXATION_VISITED_IA_COUNT) 


levels(as.factor(first4s.data$DATA_FILE))
levels(as.factor(first4s.data$condition))
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$Trial_Index_)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$Trial_Index_)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$session.num)
table(as.factor(first4s.data$DATA_FILE),
      first4s.data$session_number)
```

```{r}

first4s.data.first.fix <- first4s.data %>%
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  select(subject, session_number, IA_fam_object , condition, trial_type, trialtype_agent, agent_target_side_first, IA_FIRST_FIXATION_TIME)%>%
  pivot_wider(names_from = IA_fam_object, values_from=IA_FIRST_FIXATION_TIME)%>%
  mutate(old_goal_new_side = replace_na(old_goal_new_side, 99999999), new_goal_old_side = replace_na(new_goal_old_side, 99999999))%>%
  mutate(first_fix_old_goal=as.numeric(ifelse(old_goal_new_side<new_goal_old_side, 1, 0)),
         first_fix_old_goal_withoutNA = as.numeric(ifelse(old_goal_new_side==new_goal_old_side, NA,
                                                          ifelse(old_goal_new_side<new_goal_old_side, 1, 0))),
         first_fix_old_side=as.numeric(ifelse(old_goal_new_side>new_goal_old_side, 1, 0)),
         first_fix_old_side_withoutNA = as.numeric(ifelse(old_goal_new_side==new_goal_old_side, NA,
                                                          ifelse(old_goal_new_side>new_goal_old_side, 1, 0)))) %>%
    separate(
    agent_target_side_first,
    into = c("agent_first", "object_first", "side_first"), #shown to the dog in the first session
    sep = "_"
  ) %>%
  mutate(session_number = as.numeric(session_number), session_withinagent = as.numeric(ifelse(session_number >=3, session_number-2, session_number)))



# find the first looks to the old goal or old side. If there were first looks to one of the two IAa in both sessions we retained the first session value
first4s.data.first.fix.long <- first4s.data.first.fix %>%
  filter(first_fix_old_goal == 1 | first_fix_old_side == 1) %>%
  select(
    subject,
    trialtype_agent,
    session_withinagent,
    first_fix_old_goal,
    first_fix_old_side
  ) %>%
  pivot_longer(
    cols = c(first_fix_old_goal, first_fix_old_side),
    names_to = "IA",
    values_to = "looked"
  ) %>%
  pivot_wider(names_from = session_withinagent, values_from = looked) %>%
  rename("first.s" = "1", "second.s" = "2") %>%
  mutate(first.s = replace_na(first.s, 0), second.s = replace_na(second.s, 0))%>%
  mutate(look_overall = as.numeric(ifelse(first.s == 1, 1, ifelse(second.s == 1, 1, 0))))%>%
  filter(look_overall ==1)


table(first4s.data.first.fix.long$trialtype_agent, first4s.data.first.fix.long$IA)
mcnemar.test(table(first4s.data.first.fix.long$trialtype_agent, first4s.data.first.fix.long$IA))
```


## first fixation (binary model and no fixations are counted as 0)
```{r}
table(first4s.data.first.fix$subject, first4s.data.first.fix$session_number)
table(first4s.data.first.fix$subject, first4s.data.first.fix$session_withinagent)

first4s.data.first.fix$z.session <-
  as.vector(scale(
    as.numeric(first4s.data.first.fix$session_number),
    center = TRUE,
    scale = TRUE
  ))

first4s.data.first.fix$z.session_within <-
  as.vector(scale(
    as.numeric(first4s.data.first.fix$session_withinagent),
    center = TRUE,
    scale = TRUE
  ))

first4s.data.first.fix$trialtype_agent <-
  as.factor(first4s.data.first.fix$trialtype_agent)
levels(first4s.data.first.fix$trialtype_agent)

first4s.data.first.fix$trialtype_agent.c = as.vector(scale(
  as.numeric(
    first4s.data.first.fix$trialtype_agent == levels(as.factor(first4s.data.first.fix$trialtype_agent))[2]
  ),
  center = TRUE,
  scale = FALSE
))

first4s.data.first.fix$agent_first.c = as.vector(scale(
  as.numeric(first4s.data.first.fix$agent_first == levels(
    as.factor(first4s.data.first.fix$agent_first)
  )[2]),
  center = TRUE,
  scale = FALSE
))
levels(as.factor(first4s.data.first.fix$z.session))
hist(first4s.data.first.fix$first_fix_old_goal)
```

fit model
```{r}
mm1_first_fix=glmer(first_fix_old_goal ~ trialtype_agent+z.session_within+agent_first +
            (1+trialtype_agent.c + z.session_within|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#prune random slope structure due to convergence warning (i.e, exclude correlation between random icpt and slopes, remove random slope of session number)
mm1_first_fix2=glmer(first_fix_old_goal ~ trialtype_agent+z.session_within+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(mm1_first_fix2)
```

```{r}

mm1_first_fix2_drop1 <- drop1(mm1_first_fix2, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1)
```


Check for collinearity
```{r}
library(car)
xx=lm(first_fix_old_goal ~ trialtype_agent+z.session_within+agent_first, data=first4s.data.first.fix)
vif(xx)
```
#### Confidence intervals of the estimates
```{r}
mm1_first_fix2.ci=boot.glmm.pred(model.res=mm1_first_fix2, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)
mm1_first_fix2.ci$ci.estimates

```

#### model stability
```{r}
m.stab.mm1_first_fix2 <- glmm.model.stab(model.res = mm1_first_fix2)
m.stab.mm1_first_fix2$detailed$warnings
m.stab.mm1_first_fix2_output <- as.data.frame(round(m.stab.mm1_first_fix2$summary[, -1], 3))
m.stab.mm1_first_fix2_output 
png("graphs/mm1_first_fix2_stability_plot.png")
m.stab.plot(round(m.stab.mm1_first_fix2$summary[, -1], 3))
dev.off()
m.stab.plot(round(m.stab.mm1_first_fix2$summary[, -1], 3))
# The model appeared to be stable with respect to the fixed effects (see full_model_stability_plot).

```


#### output table


```{r}
mm1_first_fix2_output_table <-
  bind_cols(as.data.frame(summary(mm1_first_fix2)$coeff),
            mm1_first_fix2_drop1,
            mm1_first_fix2.ci$ci.estimates,
           m.stab.mm1_first_fix2_output[1:4,]) %>%
  select(
    Estimate,
    SE = `Std. Error`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`,
    min_stab = min,
    max_stab = max
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  mutate(across(.cols = c(z_wald:max_stab), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_first_fix2_output_table, file = "saves/mm1_first_fix2_output_table.csv")
```

## Plot showing the object looks in the first 4 s of the test trials (including dogs that did not look at any of the objects in a given trial)
```{r}
first4s.data.first.fix.nolooked.included.agg<-first4s.data.first.fix %>%
  group_by(trialtype_agent, session_withinagent) %>%
  summarise(first_fix_old_goal_mean = mean(first_fix_old_goal, na.rm=TRUE), 
            first_fix_old_goal_se = sd(first_fix_old_goal, na.rm=TRUE) / sqrt(length(first_fix_old_goal)),
            #number of dogs that anticipated the old goal object:
            first_fix_old_goal_sum = sum(first_fix_old_goal==1, na.rm=TRUE), 
            #number of dogs that looked at at least one of the objects:
            first_fix_old_goal_count = sum(!is.na(first_fix_old_goal)),
            first_fix_old_side_mean = mean(first_fix_old_side, na.rm=TRUE),
            first_fix_old_side_se = sd(first_fix_old_side, na.rm=TRUE) / sqrt(length(first_fix_old_side)),
            #number of dogs that anticipated the old side object:
            first_fix_old_side_sum = sum(first_fix_old_side==1, na.rm=TRUE), 
            #number of dogs that looked at at least one of the objects:
            first_fix_old_side_count = sum(!is.na(first_fix_old_side)))



al_plot_NAs_included<-ggplot(first4s.data.first.fix.nolooked.included.agg, aes(x = trialtype_agent, y = first_fix_old_goal_mean)) +
  geom_bar(stat="identity") + 
  facet_wrap(~session_withinagent)+
  ylim(0,0.5)+
  ylab("Prop. dogs fixating the old goal first")+
  xlab("Agent")+
  theme(axis.text = element_text(size = 14))+  #axis text size 
  theme(axis.title = element_text(size = 16))
```



```{r}
ggsave(filename="graphs/AL_NAs_included.png",
  plot = al_plot_NAs_included,
  width=8.5,
  height=5,
  #scale=0.55,
  dpi=1200)
```


```{r}
first4s.data.first.fix.nolooked.included.agg.long <- first4s.data.first.fix.nolooked.included.agg %>%
  select(trialtype_agent, session_withinagent, first_fix_old_goal = first_fix_old_goal_mean, first_fix_old_side = first_fix_old_side_mean) %>%
  pivot_longer(cols = c(first_fix_old_goal, first_fix_old_side), names_to = "AoI", values_to = "prop.dogs") #%>%
  #full_join(first4s.data.first.fix.nolooked.included.se.long)

al_plot_NAs_included2<-ggplot(first4s.data.first.fix.nolooked.included.agg.long, aes(x = trialtype_agent, y = prop.dogs, fill = AoI)) +
  geom_bar(stat="identity", position = position_dodge(0.9), alpha=0.8) + 
  #geom_errorbar(aes(ymin = prop.dogs - se.dogs, ymax = prop.dogs + se.dogs), width = 0.2, position = position_dodge(0.9))+
  facet_wrap(~session_withinagent)+
  ylim(0,0.5)+
  ylab("Prop. dogs")+
  xlab("Agent")+
  theme(axis.text = element_text(size = 14))+  #axis text size 
  theme(axis.title = element_text(size = 16)) +
  theme_bw() +
  scale_fill_manual(values=wes_palette("Zissou1")[c(1,4)], labels= c("old identity - 
new side","old side - 
new identity"))

al_plot_NAs_included2
```
```{r}
ggsave(filename="graphs/AL_NAs_included_AOIs.png",
  plot = al_plot_NAs_included2,
  width=8.5,
  height=5,
  scale=0.55,
  dpi=1200)
```

## first fixation (binary model and no fixations are treated as NA)
```{r}
hist(first4s.data.first.fix$first_fix_old_goal_withoutNA)
```

fit model
```{r}
mm1_first_fix_withoutNA=glmer(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session_within+agent_first +
            (1+trialtype_agent.c + z.session_within|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
#prune random slope structure due to convergence warning
mm1_first_fix_withoutNA2=glmer(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session_within+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),#removed the correlation between random intercept and random slope
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
#model has unusually large predictor estimates and variance of the random slopes
summary(mm1_first_fix_withoutNA2)
mm1_first_fix_withoutNA2_drop1<-drop1(mm1_first_fix_withoutNA2, test="Chisq")
```

pruned model (random slope of z.session removed)
```{r}
mm1_first_fix_withoutNA2_pruned=glmer(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),#removed the correlation between random intercept and random slope and the random slope of session
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
# estimates and random slope variance look better (values are smaller)
#null model because drop 1 does not converge
mm1_first_fix_withoutNA2_pruned_null=glmer(first_fix_old_goal_withoutNA ~ z.session+agent_first +
            (1|subject)+(0+trialtype_agent.c|subject),#removed the correlation between random intercept and random slope
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
anova(mm1_first_fix_withoutNA2_pruned, mm1_first_fix_withoutNA2_pruned_null, test = "Chisq")
#--> agent not significant 
summary(mm1_first_fix_withoutNA2_pruned)
drop1(mm1_first_fix_withoutNA2_pruned, test = "Chisq") #does not converge
```



Check for collinearity - model without NAs
```{r}
library(car)
xx=lm(first_fix_old_goal_withoutNA ~ trialtype_agent+z.session+agent_first, data=first4s.data.first.fix)
vif(xx)
```
#### Confidence intervals of the estimates
```{r}
mm1_first_fix2_withoutNAs.ci=boot.glmm.pred(model.res=mm1_first_fix_withoutNA2, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)
mm1_first_fix2_withoutNAs.ci$ci.estimates

```

#### model stability
```{r}
m.stab.mm1_first_fix2_withoutNAs <- glmm.model.stab(model.res = mm1_first_fix_withoutNA2)
m.stab.mm1_first_fix2_withoutNAs$detailed$warnings #models failed to converge
m.stab.mm1_first_fix2_withoutNAs_output <- as.data.frame(round(m.stab.mm1_first_fix2_withoutNAs$summary[, -1], 3))
m.stab.mm1_first_fix2_withoutNAs_output 
png("graphs/mm1_first_fix2_withoutNAs_stability_plot.png")
m.stab.plot(round(m.stab.mm1_first_fix2_withoutNAs$summary[, -1], 3))
#dev.off()
# The estimates do not look reasonable
```


#### output table


```{r}
mm1_first_fix2_withoutNAs_output_table <-
  bind_cols(as.data.frame(summary(mm1_first_fix_withoutNA2)$coeff),
            mm1_first_fix_withoutNA2_drop1,
            mm1_first_fix2_withoutNAs.ci$ci.estimates,
            m.stab.mm1_first_fix2_withoutNAs_output[1:4,]) %>%
  select(
    Estimate,
    SE = `Std. Error`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`,
    min_stab = min,
    max_stab = max
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  mutate(across(.cols = c(z_wald:max_stab), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_first_fix2_withoutNAs_output_table, file = "saves/mm1_first_fix2_withoutNAs_output_table.csv")
```

## Prepare the data for plot showing the object looks in the first 4 s of the test trials (without dogs that did not look not at any of the objects in a given trial)
```{r}
first4s.data.first.fix.agg<-first4s.data.first.fix %>%
  group_by(trialtype_agent, agent_first, session_withinagent) %>%
  summarise(first_fix_old_goal_mean = mean(first_fix_old_goal_withoutNA, na.rm=TRUE), 
            #number of dogs that anticipated the old goal object:
            first_fix_old_goal_sum = sum(first_fix_old_goal_withoutNA==1, na.rm=TRUE),   
            # number of dogs that looked at at least one of the objects:
            first_fix_old_goal_count = sum(!is.na(first_fix_old_goal_withoutNA)))
```


```{r}
first4s.data.first.fix.agg2<-first4s.data.first.fix %>%
  group_by(trialtype_agent,session_withinagent) %>%
  summarise(first_fix_old_goal_mean = mean(first_fix_old_goal_withoutNA, na.rm=TRUE), 
            first_fix_old_goal_sum = sum(first_fix_old_goal_withoutNA==1, na.rm=TRUE), #number of dogs that anticipated the old goal object
            first_fix_old_goal_count = sum(!is.na(first_fix_old_goal_withoutNA)),#number of dogs that looked at at least one of the objects
            first_fix_old_side_sum = sum(first_fix_old_side_withoutNA==0, na.rm=TRUE), #number of dogs that anticipated the old side object
            first_fix_old_side_mean=mean(first_fix_old_side_withoutNA, na.rm=TRUE),
            first_fix_old_side_count = sum(!is.na(first_fix_old_side_withoutNA)))#number of dogs that looked at at least one of the objects)

first4s.data.first.fix.nolooked.excluded.agg.long <- first4s.data.first.fix.agg2 %>%
  select(trialtype_agent, session_withinagent, first_fix_old_goal = first_fix_old_goal_mean, first_fix_old_side = first_fix_old_side_mean) %>%
  pivot_longer(cols = c(first_fix_old_goal, first_fix_old_side), names_to = "AoI", values_to = "prop.dogs") 

al_plot_NAs_excluded2<-ggplot(first4s.data.first.fix.nolooked.excluded.agg.long, aes(x = trialtype_agent, y = prop.dogs, fill = AoI)) +
  geom_bar(stat="identity", position = position_dodge(0.9), alpha=0.8) + 
  #geom_errorbar(aes(ymin = prop.dogs - se.dogs, ymax = prop.dogs + se.dogs), width = 0.2, position = position_dodge(0.9))+
  facet_wrap(~session_withinagent)+
  ylim(0,0.5)+
  ylab("Prop. dogs")+
  xlab("Agent")+
  theme(axis.text = element_text(size = 14))+  #axis text size 
  theme(axis.title = element_text(size = 16)) +
  theme_bw() +
  scale_fill_manual(values=wes_palette("Zissou1")[c(1,4)], labels= c("old identity - 
new side","old side - 
new identity"))
```

```{r}

AL_plot_NAs_excluded<-ggplot(first4s.data.first.fix.agg, aes(x = trialtype_agent, y = first_fix_old_goal_mean)) +
  geom_bar(stat="identity") + 
  facet_grid(agent_first~session_withinagent)+
  ylim(0,1)+
  ylab("Prop. dogs fixating the old goal first")+
  xlab("Agent")+
  theme(axis.text = element_text(size = 14))+  #axis text size 
  theme(axis.title = element_text(size = 16))
```

```{r}
ggsave(filename="graphs/AL_plot_NAs_excluded.png",
       plot=AL_plot_NAs_excluded,
       width=8.5,
       height=5,
       #scale=0.55,
       dpi=1200)
```




## intercept only model (to test whether the dogs have a preference for the new goal in either condition) 


```{r}
mm1_first_fix_intercept_only=glmer(first_fix_old_goal_withoutNA ~ 1+z.session+agent_first.c +
            (1|subject),
             data=first4s.data.first.fix, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mm1_first_fix_intercept_only)
```

## latency of anticipatory looks by IA
Use latency of anticipatory looks to old goal and old site as response variable and we include IA and Agent and their interaction as predictors. 

```{r}
first4s.data.latency.first.fix <- first4s.data %>%
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  select(
    subject,
    session_number,
    IA_fam_object ,
    condition,
    trial_type,
    trialtype_agent,
    agent_target_side_first,
    IA_FIRST_FIXATION_TIME,
    attention_getter_duration
  ) %>%
  mutate(IA_FIRST_FIXATION_TIME_corrected = IA_FIRST_FIXATION_TIME - attention_getter_duration)%>%
  select(-IA_FIRST_FIXATION_TIME, -attention_getter_duration)%>%
  pivot_wider(names_from = IA_fam_object, values_from = IA_FIRST_FIXATION_TIME_corrected)  %>%
      separate(
        agent_target_side_first,
        into = c("agent_first", "object_first", "side_first"),
        #shown to the dog in the first session
        sep = "_"
      ) %>%
      mutate(
        session_number = as.numeric(session_number),
        session_withinagent = as.numeric(ifelse(
          session_number >= 3, session_number - 2, session_number
        ))
      )

#numbers of trials in which the dogs looked at least at one of the two areas of interest (old goal or old side)
first4s.data.latency.first.fix %>%
  group_by(trialtype_agent, session_withinagent) %>%
  summarise(
    old_side_count = sum(!is.na(new_goal_old_side)),
    old_goal_count = sum(!is.na(old_goal_new_side)),
    both_count = sum(!is.na(new_goal_old_side) &
                       !is.na(old_goal_new_side)),
    either_count = sum(!is.na(new_goal_old_side) |
                         !is.na(old_goal_new_side))
  )
```
```{r}
first4s.data.latency.first.fix.long <- first4s.data.latency.first.fix %>%
  pivot_longer(cols = c(old_goal_new_side, new_goal_old_side), names_to = "IA", values_to = "latency") %>%
  filter(!is.na(latency))

table(first4s.data.latency.first.fix.long$subject, first4s.data.latency.first.fix.long$trialtype_agent)
table(first4s.data.latency.first.fix.long$subject, first4s.data.latency.first.fix.long$IA)

#descriptive statistics
first4s.data.latency.first.fix.long.descr<-first4s.data.latency.first.fix.long %>% 
  group_by(trialtype_agent, IA) %>% 
  summarise(mean_latency=mean(latency),
            sd_latency=sd(latency))
```


```{r}
xx.fe.re=fe.re.tab(fe.model="latency ~ IA * trialtype_agent",
re="(1|subject)", data=first4s.data.latency.first.fix.long)
xx.fe.re$summary
```

fit model (testing whether the latency to look at both AOIs is modulated by the interaction between AOI and agent)
```{r}

mm1.latency<-lme4::lmer(latency ~ IA * trialtype_agent +
            (1|subject), data=first4s.data.latency.first.fix.long, REML=FALSE)

```


```{r}
diagnostics.plot(mm1.latency, size.fac=2)

ranef.diagn.plot(mm1.latency)
```
collinearity for a model lacking the interaction
```{r}
library(car)
xx<-lm(latency ~ IA + trialtype_agent,
             data=first4s.data.latency.first.fix.long, REML=FALSE)
vif(xx)
```


Individual fixed effects
```{r}
summary(mm1.latency)
library(lmerTest)

mm1.latency.reml<-lmerTest::lmer(latency ~ IA * trialtype_agent +
            (1|subject), data=first4s.data.latency.first.fix.long, REML=TRUE)
summary(mm1.latency.reml)$coefficients
#in the inanimate condition, dogs' latency to look at the old side is lower than in the human condition

```
```{r}
first4s.data.latency.first.fix.long.agg<- first4s.data.latency.first.fix.long %>%
  rename(AoI=IA) %>% 
  group_by(trialtype_agent, AoI, subject)%>%
  summarise(mean_latency = mean(latency)*-1) #multiplied by -1 so as to have 0 as the maximum value and all other latencies as negative numbers given that these are all anticipatory looks (happening before the agents start to move)
library(wesanderson)
library(ggbeeswarm)
library(ggsignif)
library(vipor)

latency_AL_AOIs_plot<-ggplot(first4s.data.latency.first.fix.long.agg, aes(x = trialtype_agent, y = mean_latency, fill = AoI)) +
  geom_boxplot(position = position_dodge(0.8), alpha=0.8, outlier.colour = "white") +
  scale_fill_manual(values=wes_palette("Zissou1")[c(1,4)], labels= c("old identity - 
new side","old side - new identity"))+
  geom_quasirandom(method = "pseudorandom", width=0.1, dodge.width=.8, alpha=.4, size=2)+
  ylim(-3800,250)+
  ylab("Mean latency to first fixation (ms)")+
  xlab("Agent")+
  geom_signif(xmin=1.2, xmax=2.2, annotations="*", #LMM, test of fixed effect of agent (involved in interaction)
              y_position = 200, tip_length = 0.01)+
  geom_signif(xmin=1.8, xmax=2.2, annotations="*", #LMM relevelled so that inanimate is the reference category, test of fixed effect of AoI (involved in interaction)
              y_position = -50, tip_length = 0.01)+
  #geom_line(aes(group=subject), linewidth=.5, lty =2, color="grey", alpha=.9)+
  theme_bw()+
  theme(axis.text=element_text(size=12))+
  theme(axis.title=element_text(size=12))
latency_AL_AOIs_plot
```
```{r}
ggsave(filename="graphs/latency_AL_AOIs2.png",
  plot = latency_AL_AOIs_plot,
  width=10,
  height=6,
  scale=0.55,
  dpi=1200)
```

#model relevelled so that old goal is the new reference category
```{r}
first4s.data.latency.first.fix.long$IA_rel = relevel(as.factor(first4s.data.latency.first.fix.long$IA), ref = "old_goal_new_side")

levels(first4s.data.latency.first.fix.long$IA_rel)

mm1.latency.reml_rel<-lmerTest::lmer(latency ~ IA_rel * trialtype_agent +
            (1|subject), data=first4s.data.latency.first.fix.long, REML=TRUE)
summary(mm1.latency.reml_rel)
```
#model relevelled so that agent inanimate is the reference category
```{r}
first4s.data.latency.first.fix.long$trialtype_agent_rel = relevel(as.factor(first4s.data.latency.first.fix.long$trialtype_agent), ref = "inanimate")

levels(first4s.data.latency.first.fix.long$trialtype_agent_rel)

mm1.latency.reml_rel2<-lmerTest::lmer(latency ~ IA * trialtype_agent_rel +
            (1|subject), data=first4s.data.latency.first.fix.long, REML=TRUE)
summary(mm1.latency.reml_rel2)
#In the inanimate condition, dogs' latency to look at the old goal is larger than to look at the old side
```
model stability: latency 

```{r}
source("functions/glmm_stability.r")
full.stab.mm1.latency=glmm.model.stab(model.res=mm1.latency, contr=NULL,
para=F, data=NULL)

table(full.stab.mm1.latency$detailed$lme4.warnings)
table(full.stab.mm1.latency$detailed$opt.warnings)#no warnings

#table with model stability
round(full.stab.mm1.latency$summary[, -1], 3)

#plotting stability
is.re.latency=grepl(x=rownames(full.stab.mm1.latency$summary), pattern="@")
png("graphs/mm1_latency_to_first_fix_first4s_stability_plot.png")
m.stab.plot(full.stab.mm1.latency$summary[!is.re.latency, -1])#fixed effects
dev.off()
m.stab.plot(full.stab.mm1.latency$summary[is.re.latency, -1])#random effect
```
confidence intervals: latency
```{r}
source("functions/boot_glmm.r")
boot.full.mm1.latency=boot.glmm.pred(model.res=mm1.latency, excl.warnings=F,
nboots=1000, para=T, n.cores=3, resol=1000, level=0.95)
round(boot.full.mm1.latency$ci.estimates, 3)
#visualize cis
m.stab.plot(boot.full.mm1.latency$ci.estimates)
```
###Output table: latency
```{r}
model_table_latency_to_first_fix<- bind_cols(as.data.frame(summary(mm1.latency)$coefficients),
                            boot.full.mm1.latency$ci.estimates[1:4,],
                            summary(mm1.latency.reml)$coefficients,
                            full.stab.mm1.latency$summary[1:4, -1])%>% 
  select(Estimate=`Estimate...1`, SE = `Std. Error...2`, LowerCI = X2.5., UpperCI = X97.5.,t=`t value...10`, Df=`df`, p=`Pr(>|t|)`, `min`, `max`) %>%
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall=3))) %>% 
  mutate(across(.cols = c(Estimate:Df, min, max), ~ format(round(.x, 2), nsmall=2))) %>% 
# mutate(across(Chi2:p_LRT, ~replace_na(.x, "")))%>%
  mutate(p=replace(p, p==0.000, "<0.001"))

write.csv(model_table_latency_to_first_fix, file = "saves/mm1_latency_first_fix_IPfirst4s.csv")
```


## Proportion dwell time
```{r}

old_goal_object_data <- first4s.data %>%
  select(    subject,
    session_number,
    IA_fam_object ,
    condition,
    trial_type,
    trialtype_agent, 
    ia_object) %>%
  filter(IA_fam_object == "old_goal_new_side")%>%
  select(-IA_fam_object)%>%
  rename(old_goal_object = ia_object)


first4s.dwell.time <- first4s.data %>%
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  select(
    subject,
    session_number,
    IA_fam_object ,
    condition,
    trial_type,
    trialtype_agent,
    agent_target_side_first,
    IA_DWELL_TIME
  ) %>%
  pivot_wider(names_from = IA_fam_object, values_from = IA_DWELL_TIME)  %>%
      separate(
        agent_target_side_first,
        into = c("agent_first", "object_first", "side_first"),
        #shown to the dog in the first session
        sep = "_"
      ) %>%
      mutate(
        session_number = as.numeric(session_number),
        session_withinagent = as.numeric(ifelse(
          session_number >= 3, session_number - 2, session_number
        ))
      ) %>%
  mutate(prop_time_old_goal = old_goal_new_side / (old_goal_new_side + new_goal_old_side))%>%
  full_join(old_goal_object_data)
```

### beta model
```{r}

first4s.dwell.time$z.session_within<-as.vector(scale(first4s.dwell.time$session_withinagent, center = TRUE, scale=TRUE))
first4s.dwell.time$trialtype_agent<- as.factor(first4s.dwell.time$trialtype_agent)
first4s.dwell.time$trialtype_agent.c=as.vector(scale(as.numeric(first4s.dwell.time$trialtype_agent==levels(as.factor(first4s.dwell.time$trialtype_agent))[2]), center=TRUE, scale= FALSE))



first4s.dwell.time$prop_time_old_goal.scaled <- (first4s.dwell.time$prop_time_old_goal*(length(first4s.dwell.time$prop_time_old_goal) - 1) + 0.5)/length(first4s.dwell.time$prop_time_old_goal)#transform DV to exclude 0s and 1s

table(first4s.dwell.time$subject,first4s.dwell.time$prop_time_old_goal.scaled) # 5 dogs never looked at the objects during the first 4 seconds

#descriptive statistics
first4s.dwell.time.descr <- first4s.data %>%
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  select(
    subject,
    session_number,
    IA_fam_object ,
    trialtype_agent,
    IA_DWELL_TIME
  ) %>%
  pivot_wider(names_from = IA_fam_object, values_from = IA_DWELL_TIME) %>% 
   mutate(prop_time_old_goal = old_goal_new_side / (old_goal_new_side + new_goal_old_side),
          prop_time_old_side = new_goal_old_side / (new_goal_old_side + old_goal_new_side)) %>% 
  pivot_longer(cols=c(prop_time_old_goal, prop_time_old_side), names_to = "AoI", values_to = "prop_dwell_time") %>% 
  mutate(prop_dwell_time=replace_na(prop_dwell_time,0)) %>% #assign a prop. dwell time of 0 if dogs never looked into the AoIs.
  group_by(trialtype_agent,AoI) %>% 
  summarise(mean_prop_dwell_time = mean(prop_dwell_time),
            sd_prop_dwell_time = sd(prop_dwell_time))

#are there statistical differences in the prop. looking time to the old identity and old side object?
first4s.dwell.time.for.t.test <- first4s.data %>%
  filter(condition != "fam") %>% #keep only test trials
  filter(grepl("AL", IA_LABEL)) %>% #keep only AL AoIs (smaller)
  select(
    subject,
    session_number,
    IA_fam_object ,
    trialtype_agent,
    IA_DWELL_TIME
  ) %>%
  pivot_wider(names_from = IA_fam_object, values_from = IA_DWELL_TIME) %>% 
   mutate(prop_time_old_goal = old_goal_new_side / (old_goal_new_side + new_goal_old_side),
          prop_time_old_side = new_goal_old_side / (new_goal_old_side + old_goal_new_side)) %>% 
  pivot_longer(cols=c(prop_time_old_goal, prop_time_old_side), names_to = "AoI", values_to = "prop_dwell_time") %>% 
  mutate(prop_dwell_time=replace_na(prop_dwell_time,0)) %>% #assign a prop. dwell time of 0 if dogs never looked into the AoIs.
  group_by(subject, AoI) %>% 
  summarise(mean_prop_dwell_time = mean(prop_dwell_time)) %>% 
  pivot_wider(names_from = AoI, values_from = mean_prop_dwell_time)

t.test(first4s.dwell.time.for.t.test$prop_time_old_goal, first4s.dwell.time.for.t.test$prop_time_old_side, paired=TRUE, alternative = "two.sided")
```


```{r}
library(glmmTMB)
mm1.beta<-glmmTMB(prop_time_old_goal.scaled~trialtype_agent+z.session_within+agent_first+ old_goal_object + 
         #   (1+trialtype_agent.c+z.session_within+old_goal_object|subject),
            (1|subject)+
            (0+trialtype_agent.c|subject)+
            (0+z.session_within|subject),
              #(0+old_goal_object|subject),#pruned due to warning message
            family=beta_family, data=first4s.dwell.time,  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))

overdisp.test(mm1.beta)

```
```{r}
drop1_mm1.beta<-drop1(mm1.beta, test="Chisq")%>%filter(Df!="")%>%add_row(Df = rep(NA,1),  .before = 1)
drop1_mm1.beta
```

```{r}
summary(mm1.beta)
```

model stability
```{r}

mm1.beta.stab=glmmTMB.stab(model.res=mm1.beta, para=F, data=first4s.dwell.time)

mm1.beta.stab$summary

m.stab.plot(round(mm1.beta.stab$summary[1:5, -1], 3))
#is.re=grepl(x=rownames(mm1.beta.stab$summary), pattern="@")
#m.stab.plot(mm1.beta.stab$summary[!is.re, -1])
```



```{r}
library(car)
xx=lm(prop_time_old_goal.scaled~trialtype_agent+z.session_within+agent_first+ old_goal_object, data=first4s.dwell.time)
vif(xx)
```


```{r}
mm1.beta.for.ci<-glmmTMB(prop_time_old_goal.scaled~trialtype_agent+z.session_within+agent_first+ old_goal_object + 
         #   (1+trialtype_agent.c+z.session_within+old_goal_object|subject),
            (1|subject)+
            (0+trialtype_agent.c|subject)+
            (0+z.session_within|subject),
              #(0+old_goal_object|subject),#pruned due to warning message
            family=beta_family, data=first4s.dwell.time%>%filter(!is.na(prop_time_old_goal.scaled)),  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))

boot.mm1.beta<-boot.glmmtmb(mm1.beta.for.ci, data=as.data.frame(first4s.dwell.time%>%filter(!is.na(prop_time_old_goal.scaled))),
nboots=1000, para=T, n.cores="all-1", resol=100, level=0.95 )

boot.mm1.beta$ci.estimates$fe

```



### output table

```{r}
model_table_prop_dwell_time_first4s<- bind_cols(as.data.frame(summary(mm1.beta)$coefficients$cond),
                             boot.mm1.beta$ci.estimates$fe[1:5,],
                             drop1_mm1.beta,
                             mm1.beta.stab$summary[1:5, -1]) %>%
                             select(Estimate, SE = `Std. Error`, LowerCI = X2.5., UpperCI = X97.5.,  Chi2 = LRT, df = Df, p_LRT = `Pr(>Chi)`, `min`,`max`) %>%
  mutate(across(.cols = c(p_LRT), ~ format(round(.x, 3), nsmall=3))) %>% 
  mutate(across(.cols = c(Estimate:UpperCI, Chi2, min, max), ~ format(round(.x, 2), nsmall=2))) %>% 
#  mutate(across(Chi2:p_LRT, ~replace_na(.x, "")))%>%
mutate(p_LRT=replace(p_LRT, p_LRT==0.000, "<0.001"))#%>%
#mutate(p=replace(p, p==0.000, "<0.001"))

write.csv(model_table_prop_dwell_time_first4s, file = "saves/mm1_beta_prop_dwell_time_old_goal_IPfirst4s.csv")
```